```{r load, message=FALSE, echo=FALSE}
# set knitr options
source("lib_knitr.R")

# load data
load("data.rda")

# load packages and functions
library("stringr")
library("tidyverse")

source("lib_plot.R")    # colour scales and theme
```


# Taxonomic composition

## Description

Larvae were classified in `r length(unique(na.omit(d$family)))` families, `r length(unique(na.omit(d$genus)))` genera and `r length(unique(na.omit(d$species)))` identifiable species. 

CPUE is presented (still on a log10 scale, beware!) for the most common families, genera, and species below (ordered from most abundant to least abundant). Each dot represents the CPUE for one night in one site. The "violin"-shape represents the distribution of observations: the CPUE levels most commonly observed correspond to the wider regions of the shape.

```{r warning=FALSE}
# NB: do not display warnings. All warnings are about log transformation giving -Inf when using 0, which we know

# families
dd <- d0 %>% filter(!is.na(family)) %>% group_by(date, site, family) %>% summarise(cpue=sum(cpue))
dd_stat <- dd %>% group_by(family) %>% summarise(mean=mean(cpue), median=median(cpue)) %>% ungroup() %>% arrange(desc(mean)) %>% head(12)
# TODO check if there are some other interesting families beyond these 12
common_families <- dd_stat$family
dd$family <- factor(dd$family, levels=dd_stat$family)
ggplot(na.omit(dd), aes(x=family, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# genera
dd <- d0 %>% filter(!is.na(genus)) %>% group_by(date, site, family, genus) %>% summarise(cpue=sum(cpue))
dd_stat <- dd %>% group_by(family, genus) %>% summarise(mean=mean(cpue), median=median(cpue)) %>% ungroup() %>% arrange(desc(mean)) %>% head(12)
# TODO same as for families
common_genera <- dd_stat$genus
dd$genus <- factor(dd$genus, levels=dd_stat$genus)
ggplot(na.omit(dd), aes(x=genus, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# species
dd <- d0 %>% filter(!is.na(species)) %>% group_by(date, site, family, species) %>% summarise(cpue=sum(cpue))
dd_stat <- dd %>% group_by(family, species) %>% summarise(mean=mean(cpue), median=median(cpue)) %>% ungroup() %>% arrange(desc(mean))
dd_stat <- dd_stat %>% head(12)
# TODO same as for families
common_species <- dd_stat$species
dd$species <- factor(dd$species, levels=dd_stat$species)
ggplot(na.omit(dd), aes(x=species, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# add the list of common taxa to what was already in "data.rda"
save(common_families, common_genera, common_species, list=load("data.rda"), file="data.rda")
```

At the level of a any given taxon, the catch is most often < 1 fish/CARE/night/site and very variable. Despite the extreme variability, the most common taxa in CARE catches are not surprising given what we know of the coastal adult populations.


## Comparison with adult assemblages

As written above, larvae were classified in `r length(unique(na.omit(d$family)))` families, `r length(unique(na.omit(d$genus)))` genera and `r length(unique(na.omit(d$species)))` identifiable species.

```{r read_adults, results="hide"}
# read species list of north west mediterranean
library("readxl")
l <- read_excel("data/Liste especes mediterranee.xlsx", sheet=1)

# clean names written in the file
l$family <- str_trim(str_to_lower(l$family))
l$genus <- str_trim(str_to_lower(l$genus))
l$species <- str_trim(str_to_lower(l$species))

# remove undetermined species (1 atherinidae sp.)
l <- filter(l, species != "sp.")

# Join genus and species suffix in species column
l$species <- str_c(l$genus, " ", l$species)

# check that each genus has the same family
filter(summarise(group_by(l, genus), n=length(unique(family))), n>1)
filter(summarise(group_by(l, species), n=length(unique(genus))), n>1)

# remove non-demersal fish
non_demersal <- c(
"engraulidae",   # anchois
"scombridae",    # thons...
"clupeidae",     # sardines
"sphyraenidae",  # barracudas
"exocoetidae",   # poissons volants
"coryphaenidae", # coryphène
"chlorophthalmidae",
"xiphiidae",     # espadons
"istiophoridae", # marlin
"trichiuridae",  # sabres
"molidae",       # poisson lune
"regalecidae",   # regalec
"trachipteridae",
"ammodytidae",   # profond
"haemulidae",    # ??? TODO check
"macrouridae",
"aulopidae",
"petromyzonidae"
)

l$demersal <- ! l$family %in% non_demersal
l_demersal <- filter(l, demersal == TRUE)
```

According to Louisy:

- Teleost species of North Western Mediterranean are classified in `r length(unique(l$family))` families, `r length(unique(l$genus))` genera, and `r length(unique(l$species))` species. 

- Among which, dermersal species represent `r length(unique(l_demersal$family))` families, `r length(unique(l_demersal$genus))` genera, and `r length(unique(l_demersal$species))` species.

TODO check which families/genera/species are not represented in the larval catches

