```{r load, message=FALSE, echo=FALSE}
# set knitr options
source("lib_knitr.R")

# load data
load("data.rda")

# load packages and functions
library("tidyverse")

source("lib_plot.R")    # colour scales and theme
```


# Taxonomic composition

## Description

Larvae were classified in `r length(unique(na.omit(d$family)))` families, `r length(unique(na.omit(d$genus)))` genera and `r length(unique(na.omit(d$species)))` identifiable species. Among these species, 20 are catched only in one site (not necessary the same) and 8 are catched at least in 10 sites.

CPUE is presented (still on a log10 scale, beware!) for the most common families, genera, and species below (ordered from most abundant to least abundant). Each dot represents the CPUE for one night in one site. The "violin"-shape represents the distribution of observations: the CPUE levels most commonly observed correspond to the wider regions of the shape.

```{r warning=FALSE}
# NB: do not display warnings. All warnings are about log transformation giving -Inf when using 0, which we know

# families
dd <- d0 %>% filter(!is.na(family)) %>% group_by(date, site, family) %>% summarise(cpue=sum(cpue))
dd_stat <- dd %>% group_by(family) %>% summarise(mean=mean(cpue), median=median(cpue)) %>% ungroup() %>% arrange(desc(mean)) %>% head(12)
# TODO check if there are some other interesting families beyond these 12
common_families <- dd_stat$family
dd$family <- factor(dd$family, levels=dd_stat$family)
ggplot(na.omit(dd), aes(x=family, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# genera
dd <- d0 %>% filter(!is.na(genus)) %>% group_by(date, site, family, genus) %>% summarise(cpue=sum(cpue))
dd_stat <- dd %>% group_by(family, genus) %>% summarise(mean=mean(cpue), median=median(cpue)) %>% ungroup() %>% arrange(desc(mean)) %>% head(12)
# TODO same as for families
common_genera <- dd_stat$genus
dd$genus <- factor(dd$genus, levels=dd_stat$genus)
ggplot(na.omit(dd), aes(x=genus, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# species
dd <- d0 %>% filter(!is.na(species)) %>% group_by(date, site, family, species) %>% summarise(cpue=sum(cpue))
dd_stat <- dd %>% group_by(family, species) %>% summarise(mean=mean(cpue), median=median(cpue)) %>% ungroup() %>% arrange(desc(mean))
dd_stat <- dd_stat %>% head(12)
# TODO same as for families
common_species <- dd_stat$species
dd$species <- factor(dd$species, levels=dd_stat$species)
ggplot(na.omit(dd), aes(x=species, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# add the list of common taxa to what was already in "data.rda"
save(common_families, common_genera, common_species, list=load("data.rda"), file="data.rda")
```

At the level of a any given taxon, the catch is most often < 1 fish/CARE/night/site and very variable. Despite the extreme variability, the most common taxa in CARE catches are not surprising given what we know of the coastal adult populations.


## Comparison with adult assemblages
```{r warning=FALSE}
load("donees adultes.rda")
load("listesp.rda")
```

Remind : Larvae were classified in `r length(unique(na.omit(d$family)))` families, `r length(unique(na.omit(d$genus)))` genera and `r length(unique(na.omit(d$species)))` identifiable species.

Teleosts species of North Western Mediterranean are classified (according to Louisy) in `r length(unique(na.omit(l$family)))` families, `r length(unique(na.omit(l$genus)))` genera and `r length(unique(na.omit(l$species)))` species. 

Among the species most frequently observed in the adult state during counting ... To see with adults abundance data..


### Description : Abundance of adults,  example of embiez

I'm not sure it's possible to compare abundances between larvae and adults. We can compare the most abundant species between adults counting and larvae catches. But it will not be possible for all sites. The data that Anais expects are data counting from :
Carry, Carqueiranne, Embiez and Menton.

For the Embiez example , the adult data are from 2012 to 2014 (average of nb individuals / 125m²) and larval data from 2014 to 2015.

```{r warning=FALSE}


# families
aa <- a
aa_stat <- aa %>% group_by(family) %>% summarise(mean=mean(abundance))
aa_stat <- arrange(aa_stat, desc(mean))
common_family <- aa_stat$family
aa$family <- factor(aa$family, levels=aa_stat$family)
ggplot(na.omit(aa), aes(x=family, y=abundance)) +  geom_boxplot() + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# genera
aa <- a
aa_stat <- aa %>% group_by(genus) %>% summarise(mean=mean(abundance))
aa_stat <- arrange(aa_stat, desc(mean)) %>% head(12)
common_genus <- aa_stat$genus
aa$genus <- factor(aa$genus, levels=aa_stat$genus)
ggplot(na.omit(aa), aes(x=genus, y=abundance)) +  geom_boxplot() + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# species
aa <- a
aa_stat <- aa %>% group_by(species) %>% summarise(mean=mean(abundance))
aa_stat <- arrange(aa_stat, desc(mean)) %>% head(12)
common_species <- aa_stat$species
aa$species <- factor(aa$species, levels=aa_stat$species)
ggplot(na.omit(aa), aes(x=species, y=abundance)) +  geom_boxplot() + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# larvae species catched with CARE on Embiez
dd <- d0 %>% filter(!is.na(species)) %>% group_by(date, site, family, species) %>% summarise(cpue=sum(cpue))
embiez <- dd[which(dd$site=="Les Embiez"),]
dd_stat <- embiez %>% group_by(family, species) %>% summarise(mean=mean(cpue), median=median(cpue)) %>% ungroup() %>% arrange(desc(mean))
dd_stat <- dd_stat %>% head(12)
common_species <- dd_stat$species
embiez$species <- factor(embiez$species, levels=dd_stat$species)
ggplot(na.omit(embiez), aes(x=species, y=cpue)) + geom_boxplot(na.rm=T) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))
```

TODO get list of adult species and detail which are captured by CAREs or not
