---
output: 
  html_document:
    theme: null
    css: document.css
---

```{r load, message=FALSE, echo=FALSE}
# set knitr options
source("lib_knitr.R")

# load data
load("data.rda")

# load packages and functions
library("plyr")
library("dplyr")
library("stringr")
library("tidyverse")
library("readxl")
library("printr")

source("lib_plot.R")    # colour scales and theme
source("lib_data_manip.R")
```


# Taxonomic composition

## Description

### Abundance

Larvae were classified in `r length(unique(na.omit(d$family)))` families, `r length(unique(na.omit(d$genus)))` genera and `r length(unique(na.omit(d$species)))` identifiable species. 

CPUE is presented (still on a log10 scale, beware!) for the most abundant families, genera, and species below (ordered from most abundant to least abundant). Each dot represents the CPUE for one night in one site. The "violin"-shape represents the distribution of observations: the CPUE levels most commonly observed correspond to the wider regions of the shape.

```{r plots, warning=FALSE}
# NB: do not display warnings. All warnings are about log transformation giving -Inf when using 0, which we know

# families
dd <- d0 %>%
  filter(!is.na(family)) %>%
  group_by(date, site, family) %>%
  summarise(cpue=sum(cpue))
dd_stat <- dd %>%
  group_by(family) %>%
  summarise(mean=mean(cpue), median=median(cpue)) %>%
  ungroup() %>% arrange(desc(mean)) %>% head(12)
# TODO check if there are some other interesting families beyond these 12
abundant_families <- dd_stat$family
dd$family <- factor(dd$family, levels=dd_stat$family)
ggplot(na.omit(dd), aes(x=family, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# genera
dd <- d0 %>%
  filter(!is.na(genus)) %>%
  group_by(date, site, family, genus) %>%
  summarise(cpue=sum(cpue))
dd_stat <- dd %>%
  group_by(family, genus) %>%
  summarise(mean=mean(cpue), median=median(cpue)) %>%
  ungroup() %>% arrange(desc(mean)) %>% head(12)
# TODO same as for families
abundant_genera <- dd_stat$genus
dd$genus <- factor(dd$genus, levels=dd_stat$genus)
ggplot(na.omit(dd), aes(x=genus, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# species
dd <- d0 %>%
  filter(!is.na(species)) %>%
  group_by(date, site, family, species, sp) %>%
  summarise(cpue=sum(cpue))
dd_stat <- dd %>%
  group_by(family, species, sp) %>%
  summarise(mean=mean(cpue), median=median(cpue)) %>%
  ungroup() %>% arrange(desc(mean))
dd_stat <- dd_stat %>% head(12)
# TODO same as for families
abundant_species <- dd_stat$species
dd$species <- factor(dd$species, levels=dd_stat$species)
ggplot(na.omit(dd), aes(x=sp, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# add the list of common taxa to what was already in "data.rda"
save(abundant_families, abundant_genera, abundant_species, list=load("data.rda"), file="data.rda")
```

At the level of a any given taxon, the catch is most often < 1 fish/CARE/night/site and very variable. Despite the extreme variability, the most abundant taxa in CARE catches are not surprising given what we know of the coastal adult populations.


### Distribution

There are `r nrow(sites)`. Here we sort families, genera and species according to the number of sites they are caught in:

```{r freq_sp}
# compute commonality = number of sites in which a taxon is caught
commonality <- function(d, level) {
  d_common <- d %>%
    filter_(str_c("!is.na(",level,")")) %>%
    group_by_(level) %>%
    summarise(n_sites=length(unique(site))) %>%
    arrange(desc(n_sites))
  d_common[[level]] <- ordered(d_common[[level]])
  return(d_common)
}

# plot taxa in order of commonality
com_families <- commonality(d, "family")
ggplot(head(com_families, 20)) + geom_bar(aes(x=family, y=n_sites), stat="identity") + theme(axis.text.x=element_text(angle=35, hjust=1))

com_genera <- commonality(d, "genus")
ggplot(head(com_genera, 20)) + geom_bar(aes(x=genus, y=n_sites), stat="identity") + theme(axis.text.x=element_text(angle=35, hjust=1))

com_species <- commonality(d, "species")
# add abbreviated species
suppressWarnings(com_species <- left_join(com_species, select(taxo, species, sp), by="species"))
com_species$sp <- ordered(com_species$sp)
ggplot(head(com_species, 20)) + geom_bar(aes(x=sp, y=n_sites), stat="identity") + scale_y_continuous(breaks=c(0,5,10)) + theme(axis.text.x=element_text(angle=35, hjust=1))
```

These most common species are mostly the same as the most abundant ones, except for *Oblada melanura* and *Boops boops* which are abundant but present in only 3 sites.

```{r}
# keep only most common ones
common_families <- filter(com_families, n_sites > 7)$family
common_genera <- filter(com_genera, n_sites > 7)$genus
common_species <- filter(com_species, n_sites > 7)$species
save(common_families, common_genera, common_species, list=load("data.rda"), file="data.rda")
```


## Comparison with adult assemblages

As written above, larvae were classified in `r length(unique(na.omit(d$family)))` families, `r length(unique(na.omit(d$genus)))` genera and `r length(unique(na.omit(d$species)))` identifiable species.

```{r read_adults, results="hide"}
# read species list of north west mediterranean
library("readxl")
l <- read_excel("data/Liste especes mediterranee.xlsx", sheet=1)

# clean names written in the file
l$family <- str_trim(str_to_lower(l$family))
l$genus <- str_trim(str_to_lower(l$genus))
l$species <- str_trim(str_to_lower(l$species))

# remove undetermined species (1 atherinidae sp.)
l <- filter(l, species != "sp.")

# check that each genus has the same family
filter(summarise(group_by(l, genus), n=length(unique(family))), n>1)

# remove non-demersal fish
non_demersal <- c(
"engraulidae",   # anchois
"scombridae",    # thons...
"clupeidae",     # sardines
"sphyraenidae",  # barracudas
"exocoetidae",   # poissons volants
"coryphaenidae", # coryphène
"chlorophthalmidae",
"xiphiidae",     # espadons
"istiophoridae", # marlin
"trichiuridae",  # sabres
"molidae",       # poisson lune
"regalecidae",   # regalec
"trachipteridae",
"ammodytidae",   # profond  # ??? TODO check
"macrouridae",   # profond
"aulopidae",     # profond
"petromyzonidae" # eggs in river 
)

l$demersal <- ! l$family %in% non_demersal
l_demersal <- filter(l, demersal == TRUE)

l$family <- str_to_title(l$family)
l$genus <- str_to_title(l$genus)

l_demersal$family <- str_to_title(l_demersal$family)
l_demersal$genus <- str_to_title(l_demersal$genus)

# Join genus and species suffix in species column
l_demersal$species <- str_c(l_demersal$genus, " ", l_demersal$species)
```

According to Louisy:

- Teleost species of North Western Mediterranean are classified in `r length(unique(l$family))` families, `r length(unique(l$genus))` genera, and `r length(unique(l$species))` species. 

- Among which, dermersal species represent `r length(unique(l_demersal$family))` families, `r length(unique(l_demersal$genus))` genera, and `r length(unique(l_demersal$species))` species.

Comparing the list in Louisy and the list from CARE catches, the following families of demersal fish were never caught:

`r (uncaught_families <- sort(setdiff(l_demersal$family, d$family)))`.

Among families which are caught, the following genera were never caught either:

```{r}
filter(l_demersal, ! family %in% uncaught_families &! genus %in% unique(d$genus)) %>% select(family, genus) %>% unique() %>% arrange(family, genus)
```

To summarise :

- `r round((length(unique(na.omit(d$family)))*100)/length(unique(l$family)), 0)`% of families, `r round((length(unique(na.omit(d$genus)))*100)/length(unique(l$genus)), 0)`% of genera, and `r round((length(unique(na.omit(d$species)))*100)/length(unique(l$species)), 0)`% of species recorded in Louisy were catched in CAREs. 

- `r round((length(unique(na.omit(d$family)))*100)/length(unique(l_demersal$family)), 0)`% of families , `r round((length(unique(na.omit(d$genus)))*100)/length(unique(l_demersal$genus)), 0)`% of genera, and `r round((length(unique(na.omit(d$species)))*100)/length(unique(l_demersal$species)), 0)`% of species recorded as demersal fish were catched in CAREs.
