---
output: 
  html_document:
    theme: null
    css: document.css
---

```{r load, message=FALSE, echo=FALSE}
# set knitr options
source("lib_knitr.R")

# load data
load("data.rda")

# load packages and functions
library("stringr")
library("lubridate")
library("plyr")
library("tidyverse")
library("vegan")
library("FactoMineR")
library("factoextra")
library("printr")

# library("cluster")
# library("ggdendro")
# library("dendextend")
# library("NbClust")
# library("factoextra")

source("lib_plot.R")        # colour scales and theme
source("lib_data_manip.R")  # summarise tables
source("lib_factorial.R")   # extract and plot info from CAs
source("lib_quantreg.R")    # perform quantile regression
```

# Spatial distribution

## Abundances among sites

### Total community

Grouping all taxa, the catch per night is highly variable among sites, with some very extreme events yielding very high catches (NB: q25 and q75 are the quantiles at 25% and 75%, the median is the quantile at 50%; those are less sensitive than the mean to extreme values and represent the "general case" better).

```{r summary_table}
dd <- d0 %>% group_by(site, date) %>% summarise(cpue=sum(cpue))

# ggplot(dd) + geom_path(aes(x=cpue, linetype=site, color=site, y=..scaled..), stat="density", bw=0.2) + theme(legend.position=c(0.8, 0.8), legend.key.width=unit(10, "mm")) + labs(y="Scaled probability density")
# last_plot() + scale_x_continuous(trans="identity", limits=c(0,10))

dd_summary <- dd %>% sum_tbl("cpue") %>% select(-n, -n_NA)
dd_summary
```

```{r boxplot}
ggplot(dd) + geom_boxplot(aes(x=site, y=cpue)) + coord_flip()
```

```{r order}
# sort in decreasing order of catches to display stats, below
dd_summary <- arrange(dd_summary, desc(median))
```

Excluding extreme events and focusing on the more common, low (CPUE < 10), catches, allows to highlight that some sites yield better catches (`r head(dd_summary$site, 3)`) and other are consistently low (`r tail(dd_summary$site, 3)`). NB: Keeping only CPUE < 10 only eliminates `r nrow(filter(dd, cpue >= 10))` records out of `r nrow(dd)`. The black squares are quantile=0.9.

```{r zoomed_boxplot, fig.height=4}
# ggplot(dd) + geom_boxplot(aes(x=site, y=cpue, fill=..middle..), na.rm=T) + coord_flip(ylim=c(0, 10)) + scale_fill_viridis(name="median\ncpue") + theme(legend.position="bottom")
ggplot(ungroup(dd)) + geom_boxplot(aes(x=site, y=cpue), na.rm=T, outlier.size=1) + geom_quantiled(mapping=aes(x=dd$site, y=dd$cpue), quantiles=0.9, colour="black", size=2) + coord_flip(ylim=c(0, 10))
```

```{r map}
# Same information on a map with q75 in blue, median in green, and q25 in yellow.
# # add site coordinates
# dd_summary <- left_join(dd_summary, sites, by="site")
# cols <- viridis.colors(4)[-1]
# map +
#   geom_point(aes(x=lon, y=lat, size=q75), shape=16, data=dd_summary, colour=cols[1], alpha=0.7) +
#   geom_point(aes(x=lon, y=lat, size=median), shape=16, data=dd_summary, colour=cols[2], alpha=0.7) +
#   geom_point(aes(x=lon, y=lat, size=q25), shape=16, data=dd_summary, colour=cols[3], alpha=0.7) +
#   scale_size_area(max_size=20, guide="none")
```

When tested globally through a quantile based ANOVA, the lower quantiles are surprisingly the most different. The reason is that the higher quantiles (which look very different) may be more difficult to test, hence the lower significance.

```{r among_sites, echo=FALSE, }
aovq(cpue ~ site, tau=c(0.25, 0.5, 0.75, 0.9), data=dd, R=1000)
```

In terms of pairwise differences:

- q=0.25: Saint-Florent is different, higher than others
- q=0.5: Marseille and Bastia are lower
- wilcox: Marseille and Bastia are lower, Port-Cros and St-Florent are higher; Carry is often barely signif and would be similar to Marseille.
- q=0.75: ?
- q=0.9: Villefranche is higher than others; probably because it is more easily estimated (more effort).

Also, Saint Florent, Bonifacio, Villefranche, Port-Cros, Les embiez, La Ciotat, Cassis are never different.

```{r pairwise_among_sites}
pwa <- suppressMessages(pairwise.aovq(dd$cpue, dd$site, tau=c(0.25, 0.5, 0.75, 0.9), R=1000))
pww <- suppressWarnings(pairwise.wilcox.test(dd$cpue, dd$site, p.adjust.method="holm"))
pwa$wilcox <- pww
```

```{r plot_among_sites, fig.height=7}
# reformat data as data.frame
pw <- ldply(pwa, tidy)
pw$signif <- pw$p.value < 0.05
pw$group1 <- factor(pw$group1, levels=sites$site)
pw$group2 <- factor(pw$group2, levels=sites$site)
# make symmetric
pw <- bind_rows(pw, rename(pw, group1=group2, group2=group1))
# plot
ggplot(pw, aes(group1, group2)) +
  geom_point(aes(colour=p.value), na.rm=T, size=5, shape=15) +
  geom_point(colour="red", na.rm=T, size=5, shape=22, data=filter(pw, signif)) +
  facet_wrap(~.id, ncol=2) + coord_fixed() +
  scale_colour_distiller("p", palette="YlGnBu", direction=1) +
  theme(
    legend.position=c(0.75, 0.15), legend.direction="horizontal",
    # axis.text.x=element_text(angle=35, hjust=1),
    axis.text.x=element_blank(), axis.ticks.x=element_blank(),
    axis.title=element_blank()
  )
```

Overall, it seems Bonifacio, Cassis, and Carry are difficult to evaluate (low sampling effort); Bastia, Marseille, Agde, Port-Vendres, and Leucate yield low catches; Saint-Florent, Villefranche, Port-Cros, Les Embiez, La Ciotat yiled high catches.

Test whether these differences are linked with geomorphological regions:

1. Gulf of Lyon (Port-Vendres -> La Ciotat) vs Ligurian Sea (Les Embiez -> Corsica)
2. East and West of Rhône (but only 3 sites West of Rhône).

These regions are only different regarding high and extreme catches. Fish larvae abundance was higher for sites located in the Ligurian Sea than those located in the Gulf of Lions, and also higher for sites East of Rhône.

```{r abund_per_region}
# add regions for each site
dd <- left_join(dd, sites, by="site")

# test GoL vs Lig
aovq(cpue ~ topography, tau=c(0.25, 0.5, 0.75, 0.9), data=dd)
ggplot(dd) + geom_boxplot(aes(x=topography, y=cpue), na.rm = TRUE) + ylim(0, 10)

# test Rhône
aovq(cpue ~ region, tau=c(0.25, 0.5, 0.75, 0.9), data=dd)
ggplot(dd) + geom_boxplot(aes(x=region, y=cpue), na.rm = TRUE) + ylim(0, 10)
```

### Per species

The same tests can be performed per species, to detect if a given species is caught more in the West or in the East. The test was restricted to the most abundant species which are present in both locations. The following table display the mean±SD of the CPUE for each species in each region, together with the statistic and *p*-value of a quantile-based ANOVA between regions.

NB: Test done through rank procedure since bootstrap through anowar does not converge here (=sometimes low sample size, which would make any test problematic).

```{r}
# keep only abundant species, but B boops which is not caught in the West
d_abund <- filter(d0, species %in% abundant_species)

# remove the extreme values for S smaris
d_abund <- filter(d_abund, cpue < 200)

# add site characteristics
d_abund <- left_join(d_abund, sites, by="site")
```

```{r}
# topography
descriptive_stats <- d_abund %>%
  group_by(sp, topography) %>%
  summarise(stats=str_c(round(mean(cpue),2), "±", round(sd(cpue),1))) %>% 
  spread(key=topography, value=stats)
# perform test per species
tests <- d_abund %>% group_by(sp) %>% do(aovq(cpue ~ topography, tau=c(0.25, 0.5, 0.75, 0.9), data=., test="rank")$table[,c("tau", "pvalue")]) %>% spread(key=tau, value=pvalue)

# make a nice table with both
left_join(descriptive_stats, tests, by="sp") %>% ungroup()
```

```{r}
# and the corresponding plot...
ggplot(d_abund, aes(x=sp, y=cpue, colour=topography)) + geom_boxplot(na.rm=T) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=35, hjust=1))
```

```{r}
# same for region
descriptive_stats <- d_abund %>%
  group_by(sp, region) %>%
  summarise(stats=str_c(round(mean(cpue),2), "±", round(sd(cpue),1))) %>% 
  spread(key=region, value=stats)
# perform test per species
tests <- d_abund %>% group_by(sp) %>% do(aovq(cpue ~ region, tau=c(0.25, 0.5, 0.75, 0.9), data=., test="rank")$table[,c("tau", "pvalue")]) %>% spread(key=tau, value=pvalue)

# make a nice table with both
left_join(descriptive_stats, tests, by="sp") %>% ungroup()
```

```{r}
# and the corresponding plot...
ggplot(d_abund, aes(x=sp, y=cpue, colour=region)) + geom_boxplot(na.rm=T) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=35, hjust=1))
```

Overall, many differences are significant when the actual data is mostly noise above 0. Very difficult to interpret.


## Species composition among sites

### Taxonomic richness

Total number of family, genera, and species caught by sampling site.

```{r table_richness, fig.height=5}
dd_sp <- d %>%
  filter(family!="Unidentified") %>%
  group_by(site) %>%
  summarise(
    nb.family = length(unique(na.omit(family))),
    nb.genus = length(unique(na.omit(genus))),
    nb.species = length(unique(na.omit(species)))
  )
dd_sp <- arrange(dd_sp, nb.species)

# add info on effort: number of nights and of CAREs
dd <- d0 %>% group_by(site, date, n_gear) %>% summarise(cpue=sum(cpue)) %>% ungroup()
dd_night <- dd %>% group_by(site) %>% summarise(nb_night=length(unique(date)), nb_care=sum(n_gear)) %>% ungroup()
dd_sp <- left_join(dd_sp, dd_night, by=c("site"))

dd_sp
```

More species are caught in Villefranche probably because sampling is done in winter too.

Species richness vs effort (nb of nights fishes or nb of CARE deployed) curves are useful to know if this is just related to differing effort or if there is a true difference in diversity among sites. Here, the curves take a long time to saturate (>150 nights) and never really do. 

```{r specaccum, fig.height=4}
dd <- d0 %>% group_by(site, date, species) %>% summarise(cpue=sum(cpue)) %>% ungroup()
# cleanup data
dd <- filter(dd, !is.na(species))
dd <- arrange(dd, site, date)

# convert into species matrix
dw <- spread(dd, key="species", value="cpue", fill=0)

# add effort in terms of nb of CAREs (to weight by effort)
dw <- left_join(dw, effort, by=c("site", "date"))

# Species accumulation curves with CI
#
# @param see specaccum
accum_stats <- function(comm, permutations=1000, ...) {
  sp <- specaccum(comm, method="random", permutations=permutations, ...)
  stats <- adply(sp$perm, 1, function(x) {
    data.frame(
      richness=mean(x),
      sd=sd(x),
      ci.low=quantile(x, 0.025, names=FALSE),
      ci.high=quantile(x, 0.975, names=FALSE)
    )
  }, .id="effort")
  if (is.null(sp$effort)) {
    stats$effort <- sp$sites
  } else {
    stats$effort <- sp$effort
  }
  
  return(stats)
}

# compute per night
sp_per_night <- dw %>% group_by(site) %>% do(accum_stats(select(., -site, -date, -n_gear))) %>% dplyr::rename(nights=effort)
# and per CARE
# sp_per_care <- dw %>% group_by(site) %>% do(accum_stats(select(., -site, -date, -n_gear), w=.$n_gear)) %>% rename(`Nb CARE`=effort)

# and some plots
# ggplot(sp_per_night) + geom_ribbon(aes(x=nights, ymin=ci.low, ymax=ci.high, fill=site), alpha=0.2) + geom_path(aes(x=nights, y=richness, colour=site, linetype=site)) + theme(legend.key.width = unit(1.4, "cm"), legend.position="bottom")
ggplot(sp_per_night) + geom_ribbon(aes(x=nights, ymin=richness-sd, ymax=richness+sd, fill=site), alpha=0.2) + geom_path(aes(x=nights, y=richness, colour=site, linetype=site)) + theme(legend.key.width = unit(1.4, "cm"), legend.position="bottom")

# ggplot(sp_per_care) + geom_ribbon(aes(x=`Nb CARE`, ymin=ci.low, ymax=ci.high, fill=site), alpha=0.2) + geom_path(aes(x=`Nb CARE`, y=richness, colour=site, linetype=site)) + theme(legend.key.width = unit(1.4, "cm"), legend.position="bottom")
# ggplot(sp_per_care) + geom_ribbon(aes(x=`Nb CARE`, ymin=richness-sd, ymax=richness+sd, fill=site), alpha=0.2) + geom_path(aes(x=`Nb CARE`, y=richness, colour=site, linetype=site)) + theme(legend.key.width = unit(1.4, "cm"), legend.position="bottom")
```

The richness at an effort of 20 nights (largest where all stations can be tested) show that Carry is significantly less diverse than several other sites. At 50 and 150 nights, Bastia is significantly less diverse than Leucate and Villefranche. When comparable, the others are not significantly different.

```{r compare_specaccum}
# Test difference in species accumulation curves among groups at a given effort level
#
# @param comm community matrix
# @param g factor giving the group for the corresponding element of comm
# @param effort effort level(s) at which to test
# @param nperm number of permutations for the test
# @param p.adjust.method see p.adjust
pairwise.specaccum.test <- function(comm, g, effort=c(10, 50, 100), nperm=1000, p.adjust.method="holm") {
  # force grouping variable (site here) to te a factor (to keep levels consistent)
  g <- as.factor(g)
  ng <- nlevels(g)
  
  # extract richness at given efforts
  richness <- function(x, e) {
    sp <- specaccum(x, method="exact")
    sp$richness[e]
  }

  # reference (i.e. observed) richness values at these levels of effort
  cs <- split(comm, g)
  rich_ref <- laply(cs, richness, e=effort)

  # nperm permutations of sites
  rich_perm <- laply(1:nperm, function(n) {
    cs <- split(comm, sample(g))
    laply(cs, richness, e=effort)
  }, .progress="none")

  # make pairwise tests manually
  # compute all i,j under the diagonal
  ij <- expand.grid(i=1:ng, j=1:ng)
  ij <- filter(ij, i > j)
  # define a function that performs the test for a given couple of i,j indexes (i.e. couple of sites)
  testij <- function(ij) {
    i <- ij[1]
    j <- ij[2]
    # reference difference between sites
    diff_ref <- abs(diff(rich_ref[c(i,j),]))
    # permutation-based distribution of differences between sites
    diff_perm <- aaply(rich_perm[,c(i,j),], 1, function(x) {abs(diff(x))})
    # compute p-value: proportion of time in which differences obtained at random are higher than what is observed
    diffs <- rbind(diff_ref, diff_perm)
    p <- apply(diffs, 2, function(x) {sum(x[-1]>x[1])/(length(x)-1)})
    return(p)
  }
  p <- apply(ij, 1, testij)
  
  # adjust and put those p.values in a easy to display data.frame
  p <- adply(p, 1, function(pp) {
    pp <- p.adjust(pp, method=p.adjust.method, n=sum(!is.na(pp)))
    # NB: removes the NAs to compute the adjusted p.value
    d <- data.frame(ij, p.value=pp)
    return(d)
  })
  p$group1 <- levels(g)[p$i]
  p$group2 <- levels(g)[p$j]
  p$effort <- effort[p$X1]
  p <- select(p, -X1, -i, -j)

  return(p)
}

ps <- pairwise.specaccum.test(comm=select(dw, -date, -site, -n_gear), g=dw$site, effort=c(20, 50, 150))
```

```{r plot_comparison, fig.height=4}
ps$group1 <- factor(ps$group1, levels=sites$site)
ps$group2 <- factor(ps$group2, levels=sites$site)
ps$signif <- ps$p.value < 0.05
# make symmetric
ps <- bind_rows(ps, rename(ps, group1=group2, group2=group1))
# plot
ggplot(ps, aes(group1, group2)) +
  geom_point(aes(colour=p.value), na.rm=T, size=5, shape=15) +
  geom_point(colour="red", na.rm=T, size=5, shape=22, data=filter(ps, signif)) +
  facet_wrap(~effort, ncol=2) + coord_fixed() +
  scale_colour_distiller("p", palette="YlGnBu", direction=1, na.value=NA) +
  theme(
    legend.position=c(0.75, 0.15), legend.direction="horizontal",
    axis.text.x=element_text(angle=35, hjust=1),
    # axis.text.x=element_blank(), axis.ticks.x=element_blank(),
    axis.title=element_blank()
  )
```


### Site-species associations

Through a Correspondence Analysis we can study the *relative* proportions of the dominant species at each site. On the plot, sites are close to the species which were most common at this site compared to other species and other sites. In addition, sites are clustered according to a hierarchical clustering based on chi-square distances. The dendrogram cutoff as set through.... 
TODO à revoir.

So overall, the following sites are characterised by a larger proportion of the following species:

- Villefranche: *O melanuara*, *B boops*
- Port-Cros, Agde, and Port-Vendres: *M surmuletus*, *P pilicornis*
- Marseille, Carry: *D annularis*
- Leucate, Bastia: *Sarpa salpa*, *P. bogaraveo*, *P acarne*
- Bonifacio and Saint Florent: *S smaris*, *A hepsetus* and, in the same group, Les Embiez and Cassis: *C Chromis* and La Ciotat: *G mediterraneus*

Many assocations are geographic but there are some exceptions: Bastia associated with Leucate (probably because of sandy bottom and laguna in both locations); Post-Cros associated with Agde (and Port Vendres; no explanation).

```{r CA_sites, fig.height=5}
interest_species <- unique(abundant_species, common_species)
dd <- d0 %>% filter(species %in% interest_species) %>% group_by(site, species) %>% summarise(mean=mean(cpue))
ddm <- as.multivar(dd, "site")

# perform CA
ca <- CA(ddm, graph=F)

# perform hierachical clustering
dd_stand <- decostand(ddm, method = "chi.square")
dist_matrix <- vegdist(dd_stand, "euclidean")
hdw <- hclust(dist_matrix, method="ward.D2")

# Try to Find cutoff with some criterion
library("NbClust")
NbClust(data=ddm, diss=dist_matrix, distance=NULL, min.nc=2, max.nc=10, method="ward.D2", index=c("kl", "ch", "cindex", "db", "silhouette", "duda", "pseudot2", "beale", "gap"))
# TODO how to choose?

# assemble all data
ca_df <- augment(ca)
part <- data.frame(group=factor(cutree(hdw, k=6)))
part$label <- rownames(part)
ca_df <- left_join(ca_df, part, by=c("label"))

# compute variance explained
eig <- tidy(ca)
dims <- names(dat)[str_detect(names(dat), "Dim")]
labs <- str_c(dims, " (",round(eig$prop.variance[1:4]*100),"%)")

# plot result
p1 <- ggplot(ca_df, aes(x=Dim1, y=Dim2, colour=group)) +
  scale_colour_discrete(guide="none", na.value="black") +
  geom_point(size=1) +
  ggrepel::geom_text_repel(aes(label=label), size=1*2.5, segment.alpha=0.7) +
  scale_x_continuous(breaks=0) + scale_y_continuous(breaks=0) + coord_fixed() +
  labs(x=labs[1], y=labs[2])
p2 <- ggplot(ca_df, aes(x=Dim3, y=Dim4, colour=group)) +
  scale_colour_discrete(guide="none", na.value="black") +
  geom_point(size=1) +
  ggrepel::geom_text_repel(aes(label=label), size=1*2.5, segment.alpha=0.7) +
  scale_x_continuous(breaks=0) + scale_y_continuous(breaks=0) + coord_fixed() +
  labs(x=labs[3], y=labs[4])
glayout(list(p1, 1, 1), list(p2, 1, 2))
```

Using a one-way PERMANOVA based on Chi-2 distance we can explicitely test for the significance of the difference of communities between geographic or topographic regions. Species assemblages were not different between site located in the Ligurian Sea and those located in the Gulf of Lions (`topograhy`); they were according to the region (side of the Rhone).

```{r}
ddm$region <- sites$region
ddm$topography <- sites$topography
adonis(dist_matrix~region+topography, ddm, permutations = 999)$aov.tab
```


## Synchronicity among sites

### Time arrival of abundant species : Plot

```{r fig.height=8}
# mean CPUE per week, site, and species for common species
dd <- d0 %>% filter(year !=2016, species %in% abundant_species) %>% group_by(yweek, year, site, species) %>% summarise(cpue=mean(cpue))

# create a new weekly date coordinate
dd$date <- ymd(str_c(dd$year, "-01-01")) + weeks(dd$yweek)
# and a new fake one starting all the same year
dd$fake_date <- ymd("2014-01-01") + weeks(dd$yweek)

# rescale abundances per species, site and year, to get a comparable peak per year
dd <- dd %>% group_by(species, site, year) %>% mutate(scaled_cpue=cpue/max(cpue))
dd$scaled_cpue[is.na(dd$scaled_cpue)] <- 0  # when max = 0

# # remove large date jumps (i.e. add NAs at the beginning of the jump)
# dd <- arrange(dd, species, site, date) %>% ungroup()
# dd <- ddply(dd, ~species+site, function(x, large=30) {
#   large_jumps <- c(diff(x$date) > large, FALSE)
#   # repeat each line at the beginning of a large jump
#   x <- x[rep(1:nrow(x), as.numeric(large_jumps)+1),]
#   # replace the abundances at that date by NA
#   large_jumps <- c(diff(x$date) > large, FALSE)
#   x$scaled_cpue[large_jumps] <- NA
#   return(x)
# })

# rename species for display
dd$species_abbr <- abbrev_sp(dd$species, n=4)

# group sites per region
sites$region <- c(
  rep("Golfe du Lion", 3),
  rep("Marseille et environs", 4),
  rep("Var", 2),
  rep("Alpes Mar.", 1),
  rep("Corse", 3)
)

# we want to use a combination of colour and linetype to identify all sites
# ggplot2 cannot do that on its own so we need to manually compute the interaction between the two and create the appropriate scales

# prepare manual colours
colours <- scales::hue_pal(h = c(0, 360) + 15, c = 100, l = 65, h.start = 0)(length(unique(sites$region)))
sites$region_id <- as.numeric(as.factor(sites$region))
sites$colours <- colours[sites$region_id]

# prepare manual linetypes
linetypes <- c("solid", "11", "12", "21")
# linetypes <- 1:4
sites$region_change <- c(FALSE, diff(sites$region_id) != 0)
sites$linetypes <- 0
index <- 0
for (i in 1:nrow(sites)) {
  if (sites$region_change[i]) {
    index <- 1
  } else {
    index <- index + 1
  }
  sites$linetypes[i] <- index
}
sites$linetypes <- linetypes[sites$linetypes]

# compute the interaction, which will be mapped in ggplot
dd <- left_join(dd, select(sites, site, region), by="site")
dd$region_site <- interaction(dd$region, dd$site)
dd$region_site <- factor(dd$region_site, labels=sites$site)

# nice plot
ggplot(dd) +
  geom_line(aes(x=fake_date, y=scaled_cpue, colour=region_site, linetype=region_site)) +
  facet_grid(species_abbr~year)+
  scale_x_date("", date_breaks="2 month", date_labels="%b", minor_breaks=NULL) +
  theme(
    axis.text.x=element_text(angle=45, hjust=1),
    legend.position="top"
  ) +
  scale_linetype_manual("site", values=sites$linetypes) +
  scale_colour_manual("site", values=sites$colours) +
  no_y
```


### Time arrival anomaly

To investigate the time lag of larvae arrival between sites and between years, this two species were no take in count. Only the abundant and frequent species (= 10 species) were selected.  

```{r}
library(isotone)

## Remove year 2016 and select species 
res <- rbind(data.frame(species=abundant_species), data.frame(species=common_species))
interest_species <- as.character(sort(unique(res$species)))

d0$n <- d0$n_gear*d0$cpue
dd <- d0 %>% filter(year !=2016, species %in% interest_species) %>% group_by(date, yweek, yday, year, site, species) %>% summarise(cpue=sum(cpue)) %>% ungroup()


dd <- filter(dd, cpue!=0)

## Check time of arrival by species to identify winter species
ggplot(dd) +
  geom_histogram(aes(x=yweek),binwidth=1) +
  #geom_histogram(aes(x=yweek, weight=cpue),binwidth=1) +
  geom_density(aes(x=yweek, y=..count..)) +
  facet_wrap(~species, ncol=3)+geom_vline(xintercept = 0,linetype="dashed") + xlab("")

```

```{r}
#-----------

## Select winter species and choice a date to start the larvae arrival period
winter <- rbind(dd[which(dd$species=="Sarpa salpa"),], dd[which(dd$species=="Pagellus acarne"),])

winter$fake_date <- "2012-10-01"


##Modify day and week according to this new date

winter$shift_day <- round(as.double(difftime(strptime(winter$date, format = "%Y-%m-%d"),strptime(winter$fake_date, format = "%Y-%m-%d"),units="days")),0)
winter$shift_week <- round(as.double(difftime(strptime(winter$date, format = "%Y-%m-%d"),strptime(winter$fake_date, format = "%Y-%m-%d"),units="weeks")),0)

winter$season <- "2012-2013"
winter$season[which(winter$shift_day>365 & winter$shift_day<731)] <- "2013-2014"
winter$season[which(winter$shift_day>=731 & winter$shift_day<1096)] <- "2014-2015"
winter$season[which(winter$shift_day>=1096)] <- "2015-2016"

winter$shift_day <- winter$shift_day%%365
winter$shift_week <- winter$shift_week%%52


## So if we remove the season 2015-2016 (because of not data in begging 2016) we can
## consider 2012-2013 = 2012 season, 2013-2014 = 2013 season, 2014-2015 = 2014 season 
winter$fake_year <- NA
winter$fake_year[which(winter$season=="2012-2013")] <- "2013"
winter$fake_year[which(winter$season=="2013-2014")] <- "2014"
winter$fake_year[which(winter$season=="2014-2015")] <- "2015"
winter <- winter %>% filter(!is.na(fake_year)) %>% ungroup()
winter <- select(winter, year=fake_year, species, site, yweek=shift_week, cpue)

```

2012 = only species which recruit in summer
2013 = winter species (season 2012-2013) + summer species of 2013
2014 = winter species (season 2013-2014) + summer species of 2014
2015 = winter species (season 2015-2016) + summer species of 2015

```{r}
#----------------
## Summer species 
summer <- dd[-which(dd$species=="Sarpa salpa"),] 
summer <- summer[-which(summer$species=="Pagellus acarne"),]
summer <- select(summer, year, species, site, yweek,cpue)

#---------

#Paste Summer and winter species in a same data.frame
#samples <- rbind(as.data.frame(summer), as.data.frame(winter))
samples <- summer

#Compute the mean arrival week by species/sites/year
ref_detailed <- samples %>% group_by(species,site,year) %>% summarise(ref_week_det=round(weighted.mean(yweek, cpue),0))
ref_sp <- ref_detailed %>% group_by(species) %>% summarise(ref_week_sp = round(mean(ref_week_det),0))


#Join the two data.frame to comute lag time for each site
lag <- left_join(ref_detailed, ref_sp, by=c("species"))
lag$lag_week <- lag$ref_week_det - lag$ref_week_sp

```

```{r fig.height=7}
ggplot(lag) +
  geom_histogram(aes(x=lag_week),binwidth=1) +
  geom_density(aes(x=lag_week, y=..count..)) +
  facet_wrap(~site, ncol=3)+geom_vline(xintercept = 0,linetype="dashed") + xlab("")
```

Test if arrival anomaly (unit of lag_time = week) of each site is relevant with one sample wilcox.test. 

```{r}
descriptive_stats <- lag %>%
  group_by(site) %>%
  summarise(stats=str_c(round(mean(lag_week),2), "±", round(sd(lag_week),1)), median=median(lag_week))

suppressWarnings(test <- lag %>% group_by(site) %>% do(glance(wilcox.test(.$lag_week,mu = 0,alternative = "two.sided"))))

# make a nice table with both
left_join(descriptive_stats, select(test, statistic, p.value), by="site") %>% ungroup()
```

## Synchonicity among years 

Time arrival per year

```{r}

samples <- summer
# join with topography
samples <- left_join(samples, select(sites, site))

#Compute the mean arrival week by species/sites/year
ref_detailed <- samples %>% group_by(species,site,year) %>% summarise(ref_week_det=round(weighted.mean(yweek, cpue),0))
ref_sp <- ref_detailed %>% group_by(species) %>% summarise(ref_week_sp = round(mean(ref_week_det),0))


#Join the two data.frame to comute lag time for each site
lag <- left_join(ref_detailed, ref_sp, by=c("species"))
lag$lag_week <- lag$ref_week_det - lag$ref_week_sp


#  Remove 2012 because not full
lag <- filter(lag, !year == 2012)

ggplot(lag) +
  geom_histogram(aes(x=lag_week),binwidth=1) +
  geom_density(aes(x=lag_week, y=..count..)) +
  facet_wrap(~year, ncol=1)+geom_vline(xintercept = 0,linetype="dashed") + xlab("")

```

```{r}
#Compute mean and median lag 
descriptive_stats <- lag %>%
  group_by(year) %>%
  summarise(stats=str_c(round(mean(lag_week),2), "±", round(sd(lag_week),1)), median=median(lag_week)) 

suppressWarnings (test <- lag %>% group_by(year) %>% do(glance(wilcox.test(.$lag_week,mu = 0,alternative = "two.sided"))))

# make a nice table with both
left_join(descriptive_stats, select(test, statistic, p.value), by="year") %>% ungroup()
# # A tibble: 6 × 6
#     year   topography     stats median statistic p.value
#   <fctr>       <fctr>     <chr>  <dbl>     <dbl>   <dbl>
# 1   2013 Gulf of Lion  1.62±3.7      1      1550 0.18053
# 2   2013 Ligurian sea  -0.8±4.3      0      1550 0.18053
# 3   2014 Gulf of Lion  -0.7±4.5     -1      1634 0.09546
# 4   2014 Ligurian sea -0.12±4.4     -1      1634 0.09546
# 5   2015 Gulf of Lion -2.56±2.4     -2       194 0.00013
# 6   2015 Ligurian sea -1.05±4.2     -1       194 0.00013
```


Time arrival per year and topography

```{r}

samples <- summer
# join with topography
samples <- left_join(samples, select(sites, site, topography))

#Compute the mean arrival week by species/sites/year
ref_detailed <- samples %>% group_by(species,site,topography,year) %>% summarise(ref_week_det=round(weighted.mean(yweek, cpue),0))
ref_sp <- ref_detailed %>% group_by(species, topography) %>% summarise(ref_week_sp = round(mean(ref_week_det),0))


#Join the two data.frame to comute lag time for each site
lag <- left_join(ref_detailed, ref_sp, by=c("species", "topography"))
lag$lag_week <- lag$ref_week_det - lag$ref_week_sp


#  Remove 2012 because not full
lag <- filter(lag, !year == 2012)

ggplot(lag) +
  geom_histogram(aes(x=lag_week),binwidth=1) +
  geom_density(aes(x=lag_week, y=..count..)) +
  facet_wrap(year~topography, ncol=2)+geom_vline(xintercept = 0,linetype="dashed") + xlab("")

```

```{r}
#Compute mean and median lag 
descriptive_stats <- lag %>%
  group_by(year, topography) %>%
  summarise(stats=str_c(round(mean(lag_week),2), "±", round(sd(lag_week),1)), median=median(lag_week)) 

suppressWarnings (test <- lag %>% group_by(year) %>% do(glance(wilcox.test(.$lag_week,mu = 0,alternative = "two.sided"))))

# make a nice table with both
left_join(descriptive_stats, select(test, statistic, p.value), by="year") %>% ungroup()
# A tibble: 3 × 5
#     year     stats median statistic   p.value
#   <fctr>     <chr>  <dbl>     <dbl>     <dbl>
# 1   2013  0.17±4.5      1      1596 0.2598032
# 2   2014 -0.53±4.5     -1      1658 0.0297625
# 3   2015 -2.38±3.4     -2       106 0.0000083
```


2012 = only species which recruit in summer
2013 = winter species (season 2012-2013) + summer species of 2013
2014 = winter species (season 2013-2014) + summer species of 2014
2015 = winter species (season 2015-2016) + summer species of 2015



