---
output: 
  html_document:
    theme: null
    css: document.css
---

```{r load, message=FALSE, echo=FALSE}
# set knitr options
source("lib_knitr.R")

# load data
load("data.rda")

# load packages and functions
library("reshape")
library("plyr")
library("dplyr")
library("tidyverse")
library("FactoMineR")
library("printr")
library("stringr")
library("lubridate")
library("vegan")
library("ggdendro")

source("lib_plot.R")        # colour scales and theme
source("lib_data_manip.R")  # summarise tables
source("lib_factorial.R")   # extract and plot info from CAs
```

# Spatial distribution

## Abundances among sites

### Total community

Grouping all taxa, the catch per night is highly variable among sites, with some very extreme events yielding very high catches (NB: q25 and q75 are the quantiles at 25% and 75%, the median is the quantile at 50%; those are less sensitive than the mean to extreme values and represent the "general case" better).

```{r}
dd <- d0 %>% group_by(site, date) %>% summarise(cpue=sum(cpue))

# ggplot(dd) + geom_path(aes(x=cpue, linetype=site, color=site, y=..scaled..), stat="density", bw=0.2) + theme(legend.position=c(0.8, 0.8), legend.key.width=unit(10, "mm")) + labs(y="Scaled probability density")
# last_plot() + scale_x_continuous(trans="identity", limits=c(0,10))

dd_summary <- dd %>% sum_tbl("cpue") %>% select(-n, -n_NA)
dd_summary
```

```{r}
ggplot(dd) + geom_boxplot(aes(x=site, y=cpue)) + coord_flip()
```

```{r}
# compute median CPUE per site
cpue_per_site <- dd %>% group_by(site) %>% summarise(med=median(cpue)) %>% arrange(desc(med))
overall_med_cpue <- median(dd$cpue)
cpue_per_site$good <- cpue_per_site$med > overall_med_cpue
```

Excluding those extreme events and focusing on the more common, low (CPUE < 10), catches, allows to highlight that some sites yield better catches (`r head(cpue_per_site$site, 4)`) and other are consistently low (`r tail(cpue_per_site$site, 4)`). NB: Keeping only low catches eliminates `r nrow(filter(dd, cpue >= 10))` records from `r nrow(dd)`.

```{r fig.height=4}
ggplot(dd) + geom_boxplot(aes(x=site, y=cpue, fill=..middle..), na.rm=T) + coord_flip(ylim=c(0, 10)) + scale_fill_viridis(name="median\ncpue") + theme(legend.position="bottom")
```

Same on a map with q75 in blue, median in black, and q25 in grey.

```{r}
# add site coordinates
dd_summary <- left_join(dd_summary, sites, by="site")
map +
  geom_point(aes(x=lon, y=lat, size=q75), shape=16, data=dd_summary, colour="#6EB9CC", alpha=0.5) +
  geom_point(aes(x=lon, y=lat, size=median), shape=16, data=dd_summary, alpha=0.7) +
  geom_point(aes(x=lon, y=lat, size=q25), shape=16, data=dd_summary, colour="white", alpha=0.5) +
  scale_size_area(max_size=20, guide="none")
```

There does not seem to be a large scale geographical pattern to this (sites are ordered from west to east)...

Compare Gulf of Lyon vs Ligurian Sea

```{r}
# add regions for each site
dd <- left_join(dd, sites, by="site")

# test
wilcox.test(cpue ~ topography, data=dd)
ggplot(dd) + geom_boxplot(aes(x=topography, y=cpue)) + ylim(0, 10)
```

and influence of Rhône

```{r}
# test
wilcox.test(cpue ~ region, data=dd)
# plot, excluding high catches
ggplot(dd) + geom_boxplot(aes(x=region, y=cpue)) + ylim(0, 10)
```

*No statistical difference* can be detected between these regions.



## Species composition among sites


Total number of family, genus and species catched by sampling site.

```{r fig.height=5}
dd_tax <- unique(select(d, site, family, genus, species))
dd_tax <- dd_tax[-which(dd_tax$family=="Unidentified"),]

dd_sp <- ddply(dd_tax, ~ site, summarise, nb.family = length(unique(na.omit(family))),
                          nb.genus = length(unique(na.omit(genus))),
                          nb.species = length(unique(na.omit(species))))

dd_sp <- arrange(dd_sp, nb.species)
dd_sp
```

Through a Correspondence Analysis we can study the *relative* proportions of the dominant species at each site. On the plot, sites are close to the species which were most common at this site compared to other species and other sites. So overall, the following sites are characterised by a larger proportion of the following species:

- Marseille and Villefranche: *O melanuara*, *B boops*, *D annularis*
- Bonifacio and Saint Florent: *S smaris*, *A hepsetus*
- Port-Cros, Agde, and Port-Vendres: *M surmuletus*, *P pilicornis*

For the other sites, no species is particularly characteristic.

```{r fig.height=5}
dd <- d0 %>% filter(species %in% abundant_species) %>% group_by(site, species) %>% summarise(mean=mean(cpue))
ca <- CA(as.multivar(dd, "site"), graph=F)
autoplot(ca, size=1)
```

```{r fig.height=5}
## CAH clustering for sites based on species abundance
dd <- vegdist(as.multivar(dd, "site"), "bray")
hdw<-hclust(dd,method="ward.D")
plot(hdw, hang=-1)
# dend <- dendro_data(hdw, type = "rectangle")
# ggplot()+ 
#   geom_segment(data=segment(dend),aes(x = x, y = y, xend = xend, yend = yend)) + 
#   geom_text(data=label(dend), aes(x=x, y=y, label=label, hjust=1, angle=25), size=3)+
#   theme(axis.line.x=element_blank(),axis.ticks.x=element_blank(),axis.text.x=element_blank())+
#   ylab("Ultrametric distance")
```

Results of one-way PerMANOVA (to do a two-way ??) based on Bray Curtis distance matrix

```{r}
# ggplot(dd) + geom_bar(aes(x=species, y=cpue), stat="identity") + facet_grid(site~., scale="free_y") + theme(axis.text.x=element_text(angle=20, hjust=1))

## PERMANOVA
dd <- d0 %>% filter(species %in% abundant_species) %>% group_by(site, species) %>% summarise(mean=mean(cpue))
data <- cast(dd, site~species)
data.env <- left_join(data, sites, by=c("site"))
data.env <- select(data.env, site, topography, region)
data <- as.multivar(dd, "site")

as.data.frame((adonis(data~region, data.env, permutations = 999, method="bray"))$aov.tab)

(adonis(data~topography, data.env, permutations = 999, method="bray"))$aov.tab
```

TODO do the same with genera

TODO interpret in terms of spatial structure among sites (i.e. compare position of sites on CA)

## Influence of local site variables

```{r}
dd <- d0 %>% filter(!is.na(species)) %>% group_by(date, site, family, species) %>% summarise(cpue=sum(cpue))
dd_stat <- dd %>% group_by(family, species) %>% summarise(mean=mean(cpue), median=median(cpue)) %>% ungroup() %>% arrange(desc(mean))
dd_stat <- dd_stat %>% head(12)
abundant_species <- dd_stat$species

dd_test <- select(d0, site, species, cpue)
dd_test <- dd_test %>% filter(!is.na(species))
head_sp <- select(dd_stat, family, species)

dd_test <- left_join(head_sp, dd_test, by="species")

#informations of sites
dd_test <- left_join(dd_test, sites, by=c("site"))


# Barplot of mean cpue
dd_region <- dd_test %>% group_by(region, species) %>% summarise(mean = mean(cpue), sd=sd(cpue))
ggplot(dd_region, aes(x = species, y = mean))+ geom_errorbar(aes(ymin=mean, ymax=mean+sd), width=0.2)+geom_bar(stat='identity') + facet_grid(~region) +theme(axis.text.x=element_text(angle=25, hjust=1))+ylim(0,6)

dd_topography <- dd_test %>% group_by(topography, species) %>% summarise(mean = mean(cpue), sd=sd(cpue))
ggplot(dd_topography, aes(x = species, y = mean))+ geom_errorbar(aes(ymin=mean, ymax=mean+sd), width=0.2)+geom_bar(stat='identity') + facet_grid(~topography) +theme(axis.text.x=element_text(angle=25, hjust=1))+ylim(0,6)
```

Results of Kruskal-Wallis tests

```{r}
#Test differences between region but only three site in east rhone ...
# All species included
res <- dd_test %>% group_by(species, region) %>% summarise(mean =mean(cpue), sd=sd(cpue))
dd_test$region<-as.factor(dd_test$region)
test <- dd_test %>% group_by(species) %>% do(tidy(kruskal.test(.$cpue ~ .$region, data=dd_test)))
test <- as.data.frame(select(test, species, p.value))
test

#Test differences between topography 
#All species included
res <- dd_test %>% group_by(species, topography) %>% summarise(mean =mean(cpue), sd=sd(cpue))
dd_test$topography<-as.factor(dd_test$topography)
test <- dd_test %>% group_by(species) %>% do(tidy(kruskal.test(.$cpue ~ .$topography, data=dd_test)))
test <- as.data.frame(select(test, species, p.value))
test
```


TODO substrate, geomorphology, latitude, depth, etc.


## Synchronicity among sites

```{r fig.height=8}
# mean CPUE per week, site, and species for common species
dd <- d0 %>% filter(year !=2016, species %in% abundant_species) %>% group_by(yweek, year, site, species) %>% summarise(cpue=mean(cpue))

# create a new weekly date coordinate
dd$date <- ymd(str_c(dd$year, "-01-01")) + weeks(dd$yweek)
# and a new fake one starting all the same year
dd$fake_date <- ymd("2014-01-01") + weeks(dd$yweek)

# rescale abundances per species, site and year, to get a comparable peak per year
dd <- dd %>% group_by(species, site, year) %>% mutate(scaled_cpue=cpue/max(cpue))
dd$scaled_cpue[is.na(dd$scaled_cpue)] <- 0  # when max = 0

# # remove large date jumps (i.e. add NAs at the beginning of the jump)
# dd <- arrange(dd, species, site, date) %>% ungroup()
# dd <- ddply(dd, ~species+site, function(x, large=30) {
#   large_jumps <- c(diff(x$date) > large, FALSE)
#   # repeat each line at the beginning of a large jump
#   x <- x[rep(1:nrow(x), as.numeric(large_jumps)+1),]
#   # replace the abundances at that date by NA
#   large_jumps <- c(diff(x$date) > large, FALSE)
#   x$scaled_cpue[large_jumps] <- NA
#   return(x)
# })

# rename species for display
dd$species_abbr <- abbrev_sp(dd$species, n=4)

# group sites per region
sites$region <- c(
  rep("Golfe du Lion", 3),
  rep("Marseille et environs", 4),
  rep("Var", 2),
  rep("Alpes Mar.", 1),
  rep("Corse", 3)
)

# we want to use a combination of colour and linetype to identify all sites
# ggplot2 cannot do that on its own so we need to manually compute the interaction between the two and create the appropriate scales

# prepare manual colours
colours <- scales::hue_pal(h = c(0, 360) + 15, c = 100, l = 65, h.start = 0)(length(unique(sites$region)))
sites$region_id <- as.numeric(as.factor(sites$region))
sites$colours <- colours[sites$region_id]

# prepare manual linetypes
linetypes <- c("solid", "11", "12", "21")
# linetypes <- 1:4
sites$region_change <- c(FALSE, diff(sites$region_id) != 0)
sites$linetypes <- 0
index <- 0
for (i in 1:nrow(sites)) {
  if (sites$region_change[i]) {
    index <- 1
  } else {
    index <- index + 1
  }
  sites$linetypes[i] <- index
}
sites$linetypes <- linetypes[sites$linetypes]

# compute the interaction, which will be mapped in ggplot
dd <- left_join(dd, select(sites, site, region), by="site")
dd$region_site <- interaction(dd$region, dd$site)
dd$region_site <- factor(dd$region_site, labels=sites$site)

# nice plot
ggplot(dd) +
  geom_line(aes(x=fake_date, y=scaled_cpue, colour=region_site, linetype=region_site)) +
  facet_grid(species_abbr~year)+
  scale_x_date("", date_breaks="2 month", date_labels="%b", minor_breaks=NULL) +
  theme(
    axis.text.x=element_text(angle=45, hjust=1),
    legend.position="top"
  ) +
  scale_linetype_manual("site", values=sites$linetypes) +
  scale_colour_manual("site", values=sites$colours) +
  no_y
```

TODO to interpret

TODO plot, for each site and each species, the anomaly with respect to the mean arrival time of the year

## Synchronicity secund part

To investigate the time lag of larvae arrival between sites and between years, this two species were no take in count. Only the abundant and frequent species (= 10 species) were selected.  

```{r}
# mean CPUE per week, site, and species for common species
dd <- d0 %>% filter(year !=2016, species %in% abundant_species) %>% group_by(date, yweek, yday, year, site, species) %>% summarise(cpue=mean(cpue)) %>% ungroup()

# rescale abundances per species, site and year, to get a comparable peak per year
dd <- dd %>% group_by(species, site, year) %>% mutate(scaled_cpue=cpue/max(cpue))
dd$scaled_cpue[is.na(dd$scaled_cpue)] <- 0  # when max = 0
dd <- dd %>% arrange(species, year, site)
dd <- filter(dd, cpue!=0)


## Beginning of the season for winter species
winter <- rbind(dd[which(dd$species=="Sarpa salpa"),],dd[which(dd$species=="Pagellus acarne"),],
                dd[which(dd$species=="Pagellus bogaraveo"),])

winter$fake_date <- "2012-10-01"

winter$shift_day <- round(as.double(difftime(strptime(winter$date, format = "%Y-%m-%d"),strptime(winter$fake_date, format = "%Y-%m-%d"),units="days")),0)
winter$shift_week <- round(as.double(difftime(strptime(winter$date, format = "%Y-%m-%d"),strptime(winter$fake_date, format = "%Y-%m-%d"),units="weeks")),0)

winter$season <- "2012-2013"
winter$season[which(winter$shift_day>365 & winter$shift_day<731)] <- "2013-2014"
winter$season[which(winter$shift_day>731 & winter$shift_day<1096)] <- "2014-2015"
winter$season[which(winter$shift_day>1096)] <- "2015-2016"

winter$shift_day <- winter$shift_day%%365
winter$shift_week <- winter$shift_week%%52

##Check the mean arrival by season
mean <- winter %>% group_by(season, species) %>% summarise(day=mean(shift_day), week=mean(shift_week))%>%ungroup()


## So if we remove the season 2015-2016 (because of not data in begging 2016) we could be
## compare years MAYBE ???. We can consider 2012-2013 = 2012, 2013-2014 = 2013, 2014-2015 = 2014
winter$fake_year <- NA
winter$fake_year[which(winter$season=="2012-2013")] <- "2013"
winter$fake_year[which(winter$season=="2013-2014")] <- "2014"
winter$fake_year[which(winter$season=="2014-2015")] <- "2015"
winter <- winter %>% filter(!is.na(fake_year)) %>% ungroup()

syn_winter <- winter %>% group_by(fake_year, species) %>% summarise(ref_day=mean(shift_day), ref_week=mean(shift_week)) %>% ungroup()


## We have several day of arrival for each site (1256 observation in total, 1244 without end 2015) 
## and we want to compute the lag time between
## each sample with the ref time of the year, in order to plot anomalies 

## Create 2 data.frame with fake date for winter species, one for ref_time and other 
## for sample
syn_winter <- select(syn_winter,  year=fake_year, species, ref_day, ref_week)
winter <- select(winter, year=fake_year, species, site, yday=shift_day, yweek=shift_week)

##Summer species
summer <- dd[-which(dd$species=="Sarpa salpa"),] 
summer <- summer[-which(summer$species=="Pagellus acarne"),] 
summer <- summer[-which(summer$species=="Pagellus bogaraveo"),] 
summer <- select(summer, year, species, site, yday, yweek)
syn_summer <- summer %>% group_by(year, species) %>% summarise(ref_day = mean(yday), ref_week=mean(yweek))%>%ungroup() 


## All sites included, mean arrival of larvae by species(winter species + summer species) 
## by year was : 
syn_year <- rbind(syn_summer, syn_winter)
samples <- rbind(as.data.frame(summer), as.data.frame(winter))


### Compute time lag between site (by samples) for each year and each species
lag <- left_join(syn_year,samples, by = c("year","species"))
lag <- lag %>% group_by(species,year,site, ref_day, ref_week, yday, yweek) %>% summarise(lag_day = yday-ref_day, lag_week = round(yweek-ref_week))%>%ungroup()
lag <- lag %>% arrange(species,year, lag_week)

# Plot and mean lag for sites 
lag <- lag[-which(lag$species=="Boops boops"),]
lag <- lag[-which(lag$species=="Oblada melanura"),]
res_site <- lag %>% group_by(site) %>% summarise(mean = mean(lag_week), median=median(lag_week))
```

Anomaly of larvae arrival (in week) between sites

- Trends in time arrival seem to be different between sites. The first sites to recruit are Bastia, Saint-Florent, Bonifacio and Villefranche. This sites are some weeks in front of the mean date of larvae arrival. 
- Then Port-Cros, Les Embiez, La Ciotat and Cassis recruits in the same period
- Finally Marseille, Agde, Port-Vendres and Leucate recruite are the lastest site to recruit. (Carry is in front of but only 3 species, S.maris, C.chromis and D.annularis wich recruit early in year)

These differences could be due to a difference in sampling effort ?? Some sites which recruits earlier could be result of sampling only year for which time arrival of larvae was earlier than years of sampling of the other sites. 
However, there is no difference recorded between years ...
So differences between sites maybe due to the species catched in each site ?? 

```{r fig.height=7}
ggplot(lag) +
  geom_histogram(aes(x=lag_week),binwidth=1) +
  geom_density(aes(x=lag_week, y=..count..)) +
  facet_wrap(~site, ncol=3)+geom_vline(xintercept = 0,linetype="dashed") + xlab("")
```

Anomaly of larvae arrival (in week) between years 

```{r}
# Plot and mean lag
res_year <- lag %>% group_by(year) %>% summarise(mean = mean(lag_week), median = median(lag_week))
```

All sites and species included, no difference was observed between years in time arrival of larvae. However, for some species, B.Boops, S.smaris, D.annularis, O.melanura, M.surmuletus, temporal plot (cf 4.temporal Rmd) show that 2014 is in front of the other years. 


```{r fig.height=4}
ggplot(lag) +
  geom_histogram(aes(x=lag_week),binwidth=1) +
  geom_density(aes(x=lag_week, y=..count..)) +
  facet_wrap(~year, ncol=3)+geom_vline(xintercept = 0,linetype="dashed") + xlab("")
```


## Understand anomaly between sites 

Abundance distribution of species by site, to understand if time arrival anomaly between site is due to 
the species catched or realy by a lag time between sites. 

```{r}
dd_test <- select(d0, site, species, cpue)
dd_test <- dd_test %>% filter(!is.na(species))
head_sp <- select(dd_stat, family, species)
dd_test <- left_join(head_sp, dd_test, by="species")
dd_test$species <- abbrev_sp(dd_test$species, n=4)
```

```{r fig.height=8, fig.width=6}
ggplot(dd_test, aes(x=species, y=cpue)) + geom_boxplot() + scale_y_continuous(trans="log10") + facet_wrap(~site, ncol=2) + theme(axis.text.x=element_text(angle=25, hjust=1)) + ylab("log(cpue)")
```

Summary on arrival by species after corrected date for winter species

```{r}
## Mean arrival day and week by species (all years and sites included) with date corrected
## for winter species

syn <- samples %>% group_by(species) %>% summarise(ref_day = mean(yday), ref_week=mean(yweek), first_quart = quantile(yweek,0.25), third_quart = quantile(yweek,0.75)) %>% ungroup()

as.data.frame(syn)
```
