```{r options, echo=FALSE}
suppressPackageStartupMessages(library("knitr"))
opts_chunk$set(comment="#", tidy=FALSE, echo=F, fig.width=5.4, fig.height=3, dpi=100, cache=T, fig.retina=2)
options(markdown.HTML.stylesheet="~/Code/niceCSS/document.css")
```

```{r data}
load("data.rda")
```

```{r setup, cache=F}
suppressPackageStartupMessages(library("FactoMineR"))
suppressPackageStartupMessages(library("lubridate"))
suppressPackageStartupMessages(library("tidyr"))
suppressPackageStartupMessages(library("reshape2"))
suppressPackageStartupMessages(library("grid"))
suppressPackageStartupMessages(library("chroma"))
suppressPackageStartupMessages(library("broom"))
suppressPackageStartupMessages(library("stringr"))
suppressPackageStartupMessages(library("ggrepel"))
suppressPackageStartupMessages(library("ggplot2"))
suppressPackageStartupMessages(library("printr"))
suppressPackageStartupMessages(library("extrafont"))
suppressPackageStartupMessages(library("plyr"))
suppressPackageStartupMessages(library("dplyr"))

# prepare graphics
theme_set(
  theme_grey(10, "Source Sans Pro") +
  theme(
    plot.margin=unit(c(0,0,0,0), "mm"),
    axis.text = element_text(size = rel(0.9))
  )
)
scale_fill_viridis <- function(...) { scale_fill_gradientn(colors=viridis.colors(10), ...) }
scale_color_viridis <- function(...) { scale_color_gradientn(colors=viridis.colors(10), ...) }
no_y <- list(theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()), scale_y_continuous(breaks=0))

# dplyr compatibly summary function
sum_tbl <- function(x, var) {
  funs <- list(
    n=str_c("length(", var ,")"),
    n_NA=str_c("sum(is.na(", var ,"))"),
    min=str_c("min(",var,", na.rm=T)"),
    q25=str_c("stats::quantile(",var,", p=0.25, na.rm=T)"),
    median=str_c("stats::median(",var,", na.rm=T)"),
    mean=str_c("mean(",var,", na.rm=T)"),
    q75=str_c("stats::quantile(",var,", p=0.75, na.rm=T)"),
    max=str_c("max(",var,", na.rm=T)")
  )
  summarise_(x, .dots=funs)
}

# turn a data.frame like tidy object into a multivariate version
as.multivar <- function(x, x.var, y.var="species", value="mean") {
  xd <- dcast(data=x, as.formula(str_c(x.var, "~", y.var)), value.var=value, fill=0)
  rownames(xd) <- xd[,x.var]
  xd <- select_(xd, str_c("-", x.var))
  return(xd)
}


# extract and plot result from FactoMineR::CA
tidy.CA <- function(x) {
  # library("stringr")
  d <- x$eig
  names(d) <- c("eigenvalue", "prop.variance", "cum.prop.variance")
  d[,2:3] <- d[,2:3]/100
  d$term <- str_c("Dim", 1:nrow(d))
  row.names(d) <- NULL
  return(d)
}

augment.CA <- function(x, dimensions=c(1,2)) {
  # library("plyr")
  d <- ldply(c("row", "col"), function(type) {
    d <- data.frame(x[[type]]$coord)[,dimensions]
    names(d) <- str_c("Dim", dimensions)
    d$type <- type
    d$label <- row.names(d)
    d$cos2 <- rowSums(x[[type]]$cos2[,dimensions])
    return(d)
  })
}

autoplot.CA <- function(x, dimensions=c(1,2), size=1, ...) {
  # library("ggplot2")
  # library("ggrepel")
  dat <- augment(x, dimensions=dimensions)
  eig <- tidy(x)
  dims <- names(dat)[str_detect(names(dat), "Dim")]
  labs <- str_c(dims, " (",round(eig$prop.variance[dimensions]*100),"%)")
  p <- ggplot(dat, aes_string(x=dims[1], y=dims[2], colour="type")) + scale_colour_discrete(guide="none") + geom_point(size=size, ...) + geom_text_repel(aes(label=label), size=size*2.5, ...) + scale_x_continuous(breaks=0) + scale_y_continuous(breaks=0) + coord_fixed() + labs(x=labs[1], y=labs[2])
  return(p)
}

# avoid scientific notation and too many digits
options(scipen=5)
options(digits=2)
```


# MEDPLANET report, French coastline

*Jean-Olivier Irisson*  
*`r format(Sys.time(), "%Y-%m-%d")`*

## Sampling effort

Fish larvae were collected at the following sites

```{r}
map + geom_point(data=sites) + geom_text_repel(aes(label=site), data=sites, size=3)
```

At each sites, CARE light traps were deployed at several stations. The number of fish caught was summed per site and divided by the number of CARE observed to be working when collected in the morning, to compensate for sampling effort. This Catch Per Unit Effort (cpue = nb fish / CARE / night / site) is the unit used in the rest of the report. Unless otherwise stated, it includes nights with 0 catches.

Statistics on the number of CARE working per night at each site are

```{r}
# effort %>% group_by(site) %>% summarise(mean=round(mean(n_gear, na.rm=T),1), median=median(n_gear, na.rm=T), min=min(n_gear, na.rm=T), max=max(n_gear, na.rm=T)) %>% ungroup() %>% arrange(site)
effort %>% group_by(site) %>% sum_tbl("n_gear") %>% select(-n, -n_NA, -starts_with("q"))
```

Sampling occurred over the periods

```{r}
effort %>% group_by(site) %>% summarise(begin=min(date), end=max(date, na.rm=T), n_nights_sampled=length(unique(date)), `n_CARE_fished`=sum(n_gear, na.rm=T)) %>% ungroup() %>% arrange(site)
```

```{r}
ggplot(effort) + geom_point(aes(x=date, y=site), shape="|")
```


## Distribution of catches

We focus on *coastal fish* taxa only. A total of `r sum(d$n)` larvae were caught. The most commonly observed catch per night per CARE (CPUE), is 0, by far (the bars represent the number of times a given CPUE was observed in the dataset).

```{r}
dd <- d0 %>% filter(!is.na(family)) %>% group_by(date, site) %>% summarise(cpue=sum(cpue))
ggplot(dd) + geom_histogram(aes(x=cpue), binwidth=10)
```

Excluding 0, pooling all taxa together, and using log10 scaling to see low catches better (NB: notice the labels of the x-axis), the most commonly observed catches are still very low, often around 1 fish/CARE/night/site.

```{r}
ggplot(dd) + geom_histogram(aes(x=cpue), binwidth=0.1, na.rm=T) + scale_x_continuous(breaks=c(0, 1, 10, 100, 1000), trans="log10")
```

However, in terms of total number of fish supplied to the coast, its the rare, very high CPUE nights that contribute the most; the following plot is basically the previous one where each bar is multiplied by the number of larvae it corresponds to.

```{r}
ggplot(dd) + geom_histogram(aes(x=cpue, weight=cpue), binwidth=0.1, na.rm=T) + scale_x_continuous(breaks=c(0, 1, 10, 100, 1000), trans="log10") + labs(y="number of larvae")
```

```{r}
tot <- sum(dd$cpue)
dd <- dd %>% ungroup() %>% arrange(desc(cpue))
i50 <- min(which(cumsum(dd$cpue) >= tot*0.5))
percent50 <- round(i50 / nrow(dd) * 100)

# exclude nights with particularly high catches
dd <- dd[-c(1,2),]
tot <- sum(dd$cpue)
i50ex <- min(which(cumsum(dd$cpue) >= tot*0.5))
percent50ex <- round(i50ex / nrow(dd) * 100)

i90ex <- min(which(cumsum(dd$cpue) >= tot*0.9))
percent90ex <- round(i90ex / nrow(dd) * 100)
```

Overall, `r nrow(dd)` sampling "events" (one event = one night fished at one site) are recorded in the dataset. Over 50% of the `r sum(d$n)` larvae caught were caught during only `r i50` of those fishing events (`r percent50`% of the effort). Two nights in Corsica resulted in extremely high catches of *S smaris*, which mask everything else. Even excluding those, the `r percent50ex`% richest events represent >50% of the catch and the `r percent90ex`% richest yield >90% (almost all) the catch.

In summary, catches are often *low*, extremely *variable* (probably both geographically and in time) and what matters to coastal populations are the *few nights with high catches*, the rest is just negligible.


## Taxonomy

Larvae were classified in `r length(unique(na.omit(d$family)))` families, `r length(unique(na.omit(d$genus)))` genera and `r length(unique(na.omit(d$species)))` identifiable species.

CPUE is presented (still on a log10 scale, beware!) for the most common families, genera, and species below (ordered from most abundant to least abundant). Each dot represents the CPUE for one night in one site. The "violin"-shape represents the distribution of observations: the CPUE levels most commonly observed correspond to the wider regions of the shape.

```{r}
# families
dd <- d0 %>% filter(!is.na(family)) %>% group_by(date, site, family) %>% summarise(cpue=sum(cpue))
dd_stat <- dd %>% group_by(family) %>% summarise(mean=mean(cpue), median=median(cpue)) %>% ungroup() %>% arrange(desc(mean)) %>% head(12)
common_families <- dd_stat$family
dd$family <- factor(dd$family, levels=dd_stat$family)
ggplot(na.omit(dd), aes(x=family, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# genera
dd <- d0 %>% filter(!is.na(genus)) %>% group_by(date, site, family, genus) %>% summarise(cpue=sum(cpue))
dd_stat <- dd %>% group_by(family, genus) %>% summarise(mean=mean(cpue), median=median(cpue)) %>% ungroup() %>% arrange(desc(mean)) %>% head(12)
common_genera <- dd_stat$genus
dd$genus <- factor(dd$genus, levels=dd_stat$genus)
ggplot(na.omit(dd), aes(x=genus, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))

# species
dd <- d0 %>% filter(!is.na(species)) %>% group_by(date, site, family, species) %>% summarise(cpue=sum(cpue))
dd_stat <- dd %>% group_by(family, species) %>% summarise(mean=mean(cpue), median=median(cpue)) %>% ungroup() %>% arrange(desc(mean))
dd_stat <- dd_stat %>% head(12)
common_species <- dd_stat$species
dd$species <- factor(dd$species, levels=dd_stat$species)
ggplot(na.omit(dd), aes(x=species, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10") + theme(axis.text.x=element_text(angle=25, hjust=1))
```

At the level of a any given taxon, the catch is most often < 1 fish/CARE/night/site and very variable. Despite the extreme variability, the most common taxa in CARE catches are not surprising given what we know of the coastal adult populations.


## Variations among sites

Grouping all taxa, the catch per night is highly variable among sites, with some very extreme events yielding very high catches (NB: q25 and q75 are the quantiles at 25% and 75%, the median is the quantile at 50%; those are less sensitive than the mean to extreme values and represent the "general case" better).

```{r}
dd <- d0 %>% group_by(site, date) %>% summarise(cpue=sum(cpue))

# ggplot(dd) + geom_path(aes(x=cpue, linetype=site, color=site, y=..scaled..), stat="density", bw=0.2) + theme(legend.position=c(0.8, 0.8), legend.key.width=unit(10, "mm")) + labs(y="Scaled probability density")
# last_plot() + scale_x_continuous(trans="identity", limits=c(0,10))

dd %>% sum_tbl("cpue") %>% select(-n, -n_NA)
```
```{r}
ggplot(dd) + geom_boxplot(aes(x=site, y=cpue)) + coord_flip()
```

```{r}
# compute median CPUE per site
cpue_per_site <- dd %>% group_by(site) %>% summarise(med=median(cpue)) %>% arrange(desc(med))
overall_med_cpue <- median(dd$cpue)
cpue_per_site$good <- cpue_per_site$med > overall_med_cpue
```

Excluding those extreme events and focusing on the more common, low, catches, allows to highlight that some sites yield better catches (`r head(cpue_per_site$site, 4)`) and other are consistently low (`r tail(cpue_per_site$site, 4)`). There does not seem to be a large scale geographical pattern to this (sites are ordered from West to East).

```{r fig.height=4}
ggplot(dd) + geom_boxplot(aes(x=site, y=cpue, fill=..middle..), na.rm=T) + coord_flip(ylim=c(0, 10)) + scale_fill_viridis(name="median\ncpue") + theme(legend.position="bottom")
```

Through a Correspondence Analysis we can study the *relative* proportions of the dominant species at each site. On the plot, sites are close to the species which were most common at this site compared to other species and other sites. So overall, the following sites are characterised by a larger proportion of the following species:

- Marseille and Villefranche: *O melanuara*, *B boops*, *D annularis*
- Bonifacio and Saint Florent: *S smaris*, *A hepsetus*
- Port-Cros, Agde, and Port-Vendres: *M surmuletus*, *P pilicornis*

For the other sites, no species is particularly characteristic.

```{r fig.height=5}
dd <- d0 %>% filter(species %in% common_species) %>% group_by(site, species) %>% summarise(mean=mean(cpue))
ca <- CA(as.multivar(dd, "site"), graph=F)
autoplot(ca, size=1)

# ggplot(dd) + geom_bar(aes(x=species, y=cpue), stat="identity") + facet_grid(site~., scale="free_y") + theme(axis.text.x=element_text(angle=20, hjust=1))
```



## Variations among years

The effort varied quite a lot among years.

```{r}
effort %>% mutate(year=lubridate::year(date)) %>% filter(year != 2016) %>% group_by(year) %>% summarise(n_nights_sampled=length(unique(date)), n_CARE_fished=sum(n_gear))
```

In total CPUE, pooling all taxa together, the catch seems similar over the years, with possibly a shift towards higher catches in 2014 and 2015. 

```{r}
dd <- d0 %>% filter(year != 2016) %>% group_by(date, year, site) %>% summarise(cpue=sum(cpue))

dd %>% group_by(year) %>% sum_tbl("cpue") %>% select(-n, -n_NA)

ggplot(dd, aes(x=year, y=cpue)) + geom_violin(na.rm=T) + geom_point(alpha=0.2, size=1) + scale_y_continuous(trans="log10")
```

When inspecting the dynamics per species the characteristic year-species couples are:

- 2012: *A hepsetus*
- 2013: *S smaris*
- 2014: *B boops*, *D annularis*, *O melanuara*

```{r fig.height=5}
dd <- d0 %>% filter(year != 2016, species %in% common_species) %>% group_by(year, species) %>% summarise(tot=sum(cpue), mean=mean(cpue), sd=sd(cpue), median=median(cpue), mad=mad(cpue))

# relative abundance
# ggplot(dd) + geom_bar(aes(x=year, y=mean), stat="identity") + facet_wrap(~species, scales="free_y") + theme(axis.text.x=element_text(angle=30, hjust=1))

# just make the plot cleaner by pre-computing the relative abundances
ddr <- dd %>% group_by(species) %>% mutate(mean=mean/max(mean))
ggplot(ddr) +
  geom_bar(aes(x=year, y=mean), stat="identity", width=0.75) +
  facet_wrap(~species, ncol=3) +
  scale_y_continuous("relative cpue", breaks=c(0, 0.5, 1)) +
  theme(axis.text.x=element_text(angle=30, hjust=1), axis.text.y=element_blank(), axis.ticks.y=element_blank())

ca <- CA(as.multivar(dd, x.var="year"), graph=F)
autoplot(ca)
```


## Seasonal variations, within years

Unsurprisingly, the catch is very seasonal, with higher catches from May to mid-September and another small increase around November. NB: the two points with very high catches of *S smaris* occurred in June; they were removed from the plot.

```{r}
dd <- d0 %>% group_by(date, yday, year, site) %>% summarise(cpue=sum(cpue))
# remove high catches of S smaris
dd <- filter(dd, cpue < 110)
# create a fake date coordinate
dd$fake_date <- ymd("2014-01-01") + days(dd$yday)

ggplot(dd, aes(x=fake_date, y=cpue)) + geom_point(size=1, alpha=0.5) + geom_smooth(size=0.5, color="#F55D5B", fill="#F55D5B") + scale_x_datetime("month", date_breaks="1 month", date_labels="%b")
```

At species level, the time of arrival has consistent seasonality across years but can vary of a few weeks from year to year.

```{r fig.height=7}
dd <- d0 %>% filter(year != 2016, species %in% common_species) %>% group_by(yweek, year, species) %>% summarise(cpue=mean(cpue))

# create a fake date coordinate
dd$fake_date <- ymd("2014-01-01") + weeks(dd$yweek)

# rescale per species and year
dd <- dd %>% group_by(species, year) %>% mutate(scaled_cpue=cpue/max(cpue))
dd$scaled_cpue[is.na(dd$scaled_cpue)] <- 0  # when max = 0

# order species by average/first date of arrival
order <- dd %>%
  group_by(species) %>%
  filter(cpue > median(cpue)) %>% 
  summarise(avg_date=weighted.mean(yweek, scaled_cpue), first_date=min(yweek)) %>%
  arrange(first_date)

# rename species for display
abbrev_sp <- function(x, n=4) {
  bits <- str_split_fixed(x, fixed(" "), 2)
  g <- str_sub(bits[,1],1,1)
  s <- str_sub(bits[,2],1,n)
  str_c(g, ". ", s)
}
order$species_abbr <- abbrev_sp(order$species, n=5)

# force order by time of arrival
dd$species_abbr <- factor(dd$species, levels=order$species, labels=order$species_abbr)

ggplot(dd) +
  geom_path(aes(x=fake_date, y=scaled_cpue, colour=year)) +
  facet_grid(species_abbr~.) +
  scale_x_datetime(date_breaks="1 month", date_labels="%b", minor_breaks=NULL) +
  labs(y="cpue (scaled per species and year)") + no_y
```


## Lunar phase variations

```{r results="hide"}
suppressPackageStartupMessages(library("oce", quietly=TRUE))
# get all dates
dates <- sort(unique(d0$date))

# for each date, compute the proportion of the night when the moon is up
moon <- ldply(dates, function(x) {
  # x <- dates[8]

  # go back from noon the previous day, to noon on the day of collection
  interval <- 0.1
  xr <- x - days(1) + seq(12, 36, by=interval)*3600

  # compute position of sun and moon
  sun_alt <- sunAngle(xr, lon=7.31, lat=43.68)$altitude
  moon <- moonAngle(xr, lon=7.31, lat=43.68)
  moon_alt <- moon$altitude

  # detect when sun is down and moon is up
  moon_visible <- (sun_alt < 0) & (moon_alt > 0)
  # compute the number of hours this is the case
  moon_time <- sum(moon_visible) * interval
  
  # compute a moon phase index, based on the proportion illuminated
  # 0 : new moon
  # 1 : next full moon
  # -1 : previous full moon
  # negative : descending
  # positive : ascending
  moon_phase <- mean(moon$illuminatedFraction[sun_alt < 0]) * sign(diff(moon$illuminatedFraction[c(1, length(xr))]))
  
  # compute average "moonlight" indexes during that time
  if (moon_time > 0) {
    # visible fraction
    moon_index1 <- mean(moon$illuminatedFraction[moon_visible]) / moon_time
  
    # fraction * diameter * altitude
    moon_index2 <- sum(moon$illuminatedFraction[moon_visible] * moon$diameter[moon_visible] * (moon$altitude[moon_visible]/90))
  } else {
    moon_index1 <- moon_index2 <- 0
  }
  
  data.frame(date=x, moon_time, moon_phase, moon_index1, moon_index2)
}, .progress="none")

# add that to the data
d0m <- left_join(d0, moon, by="date")
dm <- left_join(d, moon, by="date")

# # interpolate times of rise and set of celestial bodies
# rise_set <- function(time, altitude) {
#   # time <- ymd("2016-01-01") + hours(0:(24*3))
#   # altitude <- sunAngle(time, lon=7.31, lat=43.68)$altitude
#   xn <- as.numeric(time)
#
#   # detect sign changes in altitude
#   sign_change <- diff(sign(altitude))
#   i <- which(sign_change != 0)
#
#   # interpolate time when body crosses altitude 0
#   d <- data.frame(x1=xn[i], x2=xn[i+1], y1=altitude[i], y2=altitude[i+1])
#   times <- apply(d, 1, function(X) {approx(X[3:4], X[1:2], value)$y})
#   times <- as.POSIXct(times, tz=tz(time), origin="1970-01-01")
#
#   # determine crossing direction
#   direction <- c("set", "rise")[as.numeric(sign_change[i]>0)+1]
#
#   return(data.frame(time=times, direction))
# }
```

If we look at the total number of larvae caught, a large proportion of it is caught around the new moon, as shown below for common species. However the second plot shows that there is still considerable variability in CPUE, even around the new moon, and that it mostly masks the moon-phase signal (the red lines are polynomial regressions of the average CPUE; most of which do not highlight a moon phase influence).


```{r fig.height=5}
moon_x <- list(scale_x_continuous(breaks=c(-1, 0, 1), labels=c("\U0099", "\U0098", "\U0099")), theme(axis.text.x=element_text(family="Wingdings 2"), axis.title.x=element_blank()))

ggplot(filter(d0m, species %in% common_species)) + geom_histogram(aes(x=moon_phase, weight=cpue, y=..ncount..), binwidth=0.05, fill="grey30", colour=NA) + facet_wrap(~species, ncol=3) + moon_x + no_y + labs(y="total number of larvae (scaled per species)")

# # density
# ggplot(filter(d0m, species %in% common_species)) + geom_density(aes(x=moon_phase, weight=cpue, y=..scaled..), bw=0.05, fill="grey30", colour=NA) + facet_wrap(~species, ncol=3) + moon_x + no_y
#
# # density logged
# ggplot(filter(d0m, species %in% common_species)) + geom_density(aes(x=moon_phase, weight=log1p(cpue), y=..scaled..), bw=0.05, fill="grey30", colour=NA) + facet_wrap(~species, ncol=3) + moon_x + no_y
# # does not change much

ggplot(filter(d0m, species %in% common_species), aes(x=moon_phase, y=cpue)) + geom_point(size=1, alpha=0.5) + geom_smooth(se=F, colour="#F55D5B", size=0.5, data=filter(dm, species %in% common_species)) + facet_wrap(~species, ncol=3, scales="free_y") + moon_x + no_y + labs(y="cpue (scaled per species)")

# # log10 transformed
# ggplot(filter(d0m, species %in% common_species), aes(x=moon_phase, y=cpue)) + geom_point(size=1, alpha=0.5) + geom_smooth(se=F, colour="red", size=0.5) + facet_wrap(~species, ncol=3) + scale_y_log10() + moon_x
```

Even when we reduce the data to summer months (May to August, included), during which the catches are better overall, the moon phase signal is still mostly indiscernible in instantaneous CPUE.

```{r fig.height=5}
# reduce to summer months
in_summer <- function(x) {
  month(x) %in% 5:8
}
ggplot(filter(d0m, species %in% common_species, in_summer(date)), aes(x=moon_phase, y=cpue)) + geom_point(size=1, alpha=0.5) + geom_smooth(se=F, colour="#F55D5B", size=0.5, data=filter(dm, species %in% common_species)) + facet_wrap(~species, ncol=3, scales="free_y") + moon_x + no_y + labs(y="summer cpue (scaled per species)")
```

Another way to look at it is to regress CPUE on moon phase. Using quantile regression, we can focus on the common scenario (median, quantile=0.5), or the higher catches (quantiles > 0.5). We see that the median catch is basically 0 everywhere and that only the 90th percentile (quantile=0.9) starts to respond to moon phase.

```{r fig.height=4}
suppressPackageStartupMessages(library("quantreg", quietly=TRUE))

# select data: common species, in summer, without extreme Spicara
dd <- filter(d0m, species %in% common_species, in_summer(date), cpue < 250)

# only consider distance to the new moon, irrespective of direction
dd$moon_phase <- abs(dd$moon_phase)

# # plot according to various moon related variables
# ggplot(dd) + geom_point(aes(x=moon_time, y=cpue))
# ggplot(dd) + geom_point(aes(x=moon_phase, y=cpue))
# ggplot(dd) + geom_point(aes(x=moon_index1, y=cpue))
# ggplot(dd) + geom_point(aes(x=moon_index2, y=cpue))

# fit and predict quantile regression models
x <- seq(0, 1, length.out=100)
new <- data.frame(moon_phase=x, moon_index1=x, moon_index2=x)

quantiles <- c(0.5, 0.75, 0.9, 0.95, 0.99)
qpred <- ldply(quantiles, function(q) {
  # quantile regression on log(n+1)
  m <- rq(log1p(cpue) ~ moon_phase, tau=q, data=dd)

  # predict regression curve
  pred <- tryCatch(
    predict(m, newdata=new, interval="confidence"),
    error=function(x) {return(data.frame(fit=rep(0, times=100), lower=0, higher=0))}
  )
  # switch back to cpue
  pred <- expm1(pred)
  
  # add "metadata"
  pred <- cbind(new, pred)
  pred$quantile <- q

  return(pred)
})
qpred$quantile <- factor(qpred$quantile)

# plot the result
ggplot(mapping=aes(x=moon_phase)) +
  geom_point(aes(y=cpue), data=dd, alpha=0.5) +
  geom_ribbon(aes(ymin=lower, ymax=higher, fill=quantile), data=qpred, alpha=0.3) +
  geom_path(aes(y=fit, colour=quantile), data=qpred) +
  # scale_colour_brewer(palette="Reds") + scale_fill_brewer(palette="Reds") +
  scale_y_continuous("cpue (in summer, log(n+1) transformed)", trans="log1p") +
  scale_x_continuous("moon phase", breaks=c(0, 0.5, 1), labels=c("\U0098", "\U00BA", "\U0099")) +
  theme(axis.text.x=element_text(family="Wingdings 2"))
```

So, overall, high catches almost always occur around the new moon but not all new moons yield high catches (actually, most do not), even in the good season.


## Temporal scales of variability

The previous analyses are done at various temporal scales. Now we can analyse the temporal scales of variability through a variogram. This computes the difference in CPUE between two successive days (delay = 1 day), two days on successive weeks (delay = 7 days), etc. to test whether the catches are more similar on successive days (autocorrelation) or on the same day in successive months (lunar effect), etc.

This is done within each site, for all sampled nights, for the most common species (for which catches are relatively steady). The results below are for all pairwise differences (top) and some summaries --quartiles, median and mean-- (red lines on top, bottom).

```{r}
dd <- d0 %>% filter(species %in% common_species) %>% group_by(days_since_start, site) %>% summarise(cpue=sum(cpue)) %>% group_by(site) %>% arrange(days_since_start)

summary(dd$cpue)
#  Min. 1st Qu.  Median    Mean 3rd Qu.    Max.
# 0.000   0.000   0.375   3.736   1.625 899.000

# compute variogram cloud (all pairwise differences)
vario_cloud <- ddply(dd, ~site, function(x) {
  delay <- dist(x$days_since_start)
  # diff <- dist(log1p(x$cpue))
  diff <- dist(x$cpue)
  d <- data.frame(delay=as.numeric(delay), diff=as.numeric(diff))
})
vario_cloud <- filter(vario_cloud, delay < (365*2+10))

# compute average variogram on bins with an ~ equal number of observations
vario <- vario_cloud
breaks <- sort(unique(c(2, 5, quantile(vario$delay, seq(0,1,length.out=30)))))
vario$bins <- cut(vario$delay, breaks, right=F, include.lowest=T)

diffs <- vario %>% group_by(bins) %>% sum_tbl("diff")
delay <- vario %>% group_by(bins) %>% summarise(delay=min(delay))
vario <- full_join(delay, diffs, by="bins")

x_scale <- list(
  scale_x_continuous(
    limits=c(0, 715),
    breaks=c(1, 30, 30*3, 30*6, 365, 365*2),
    labels=c("1 day", "1 month", "3 months", "6 months", "1 year", "2 years"),
    minor=NULL
  ),
  theme(axis.text.x=element_text(angle=35, hjust=1))
)

last_point <- vario[nrow(vario),] %>% select(delay, median, mean) %>% gather(key=label, value=diff, -delay)
ggplot(mapping=aes(x=delay, y=diff)) +
  geom_point(data=vario_cloud, alpha=0.02, shape=16, na.rm=T) +
  geom_path(aes(y=median), data=vario, colour="#F55D5B") +
  geom_path(aes(y=mean), data=vario, colour="#F55D5B", linetype="dashed") +
  geom_label(aes(y=diff, label=label), data=last_point, hjust=0, nudge_x=5, size=3, colour="#F55D5B") +
  x_scale + scale_y_continuous(limits=c(0,100), trans="log1p") +
  labs(x="delay (days)", y="difference in CPUE (log(n+1) scale)")

last_point <- vario[nrow(vario),] %>% select(delay, q25, median, mean, q75) %>% gather(key=label, value=diff, -delay)
ggplot(data=vario, mapping=aes(x=delay)) +
  geom_ribbon(aes(ymin=q25, ymax=q75), alpha=0.5) +
  geom_path(aes(y=median)) +
  geom_path(aes(y=mean), linetype="dashed") +
  geom_text(aes(y=diff, label=label), data=last_point, hjust=0, nudge_x=5, size=3) +
  x_scale + scale_y_continuous(limits=c(0,NA)) +
  labs(x="delay (days)", y="difference in CPUE")
```

These plots can be interpreted as follows:

- from one day to the next (left extreme of the plot), the median difference in CPUE is about 0.5 to 1 fish/CARE/night/site, when the median CPUE is `r median(dd$cpue)`. In other words, the variability is considerably greater than the observation, meaning that, in the vast majority of cases, how much you catch today does not allow to predict how much you will catch tomorrow.
- variance is even higher from one lunar cycle to the next (one month delay).
- a clear ~ 6 months long oscillation is visible, which probably matches the seasonal component highlighted above:
    - high catches in spring-summer and somewhat higher catches in November (6 months between May and Nov),
    - and, more importantly, steady, low catches in Sept-Oct and from Jan to April (6 months between Sept and March).

Overall, catches are as variable in the short term (day to day) as in the long term (month to month or year to year). This, combined with the importance of the large peaks of settlement for the total catch, has important consequences for *sampling strategies*. Indeed, sampling too sparsely in time will likely lead to miss large peaks and change the interpretation of the data completely. Actually, from this analysis, the only safe sampling strategy would seem to be to sample *every single day*.

## Synchronicity among sites

todo

# Note on data format

To prepare this report, the data from several sources had to be homogenised. This was both long and cumbersome. A good organisation for the data resulting from CARE traps would be the following.

A table `effort` in which information is recorded every time a CARE is retrieved from the water. At minima, it should contain

```{r}
suppressPackageStartupMessages(library("readxl", quietly=TRUE))
read_excel("sample_record/effort.xlsx", sheet=1, na="NA")
```

where

- `date_out` : date as which the CARE was retrieved, in `YYYY-MM-DD` format (ISO 8601).
- `station` : name of site; general location where CAREs are set.
- `station` : name of station; a station is a location within a site; stations would typically be a few hundred meters apart and data from various stations will likely be grouped together in the analysis.
- `lat`, `lon` : location of station; if possible in decimal format, alternatively in a consistent degreesºminutes'seconds" format.
- `n_gear_in`: number of CARE put in the water at the station.
- `n_gear_ok`: number of CARE retrieved and checked to be working (light on, collector shut, etc.); this is what will be used to compute fishing effort.
- `comments`: well, comments `;-)`

Additional information such as bottom depth, weather, time of retrieval etc. can also be recorded in this table. If some information is often repeated (such as coordinates, if the stations are always the same), those can be stored in a different table and linked to this one.


A table `catches` in which information is recorded only when a fish larva is caught. Again, a minimal table would contain

```{r}
read_excel("sample_record/catches.xlsx", sheet=1, na="NA")

```

where 

- `date_out`, `site`, and `station` allow to match up with the records in `effort`.
- `family`, `genus`, `species`: taxonomic denomination of the larvae. When an element cannot be determined:
    - either confusion between several taxa is possible (several genera, or several species --usually noted spp.), in that case leave it blank
    - or the larvae can be determined as all belonging to one taxon (one genus, one species) but this taxon cannot be identified, in that case give it a generic name such as `gen1`, `sp1`, `sp_dark`, and, if possible, document it in the comments (see example above).
- `n`: number of larvae.
- `comments`: well, still comments, but on the fish now.

Additional information such as developmental stage, size, etc. can be recorded here.

Zero catches do *not* need to be recorded (and should *not* be) because they can be deduced from the matchup with `effort`. For example, one CARE was put in the water at site A on 2012-10-24 but nothing is recorded in `catches`, meaning that nothing was caught.





