---
output: 
  html_document:
    theme: null
    css: document.css
---

```{r load, message=FALSE, echo=FALSE}
# set knitr options
source("lib_knitr.R")
library("printr")
```

# Note on data format

The data from several sources often has to be homogenised. This is both long and cumbersome. A good organisation for the data resulting from CARE traps would be the following.

A table `effort` in which information is recorded every time a CARE is retrieved from the water. At minima, it should contain

```{r}
suppressPackageStartupMessages(library("readxl", quietly=TRUE))
read_excel("sample_record/effort.xlsx", sheet=1, na="NA")
```

where

- `date_out` : date as which the CARE was retrieved, in `YYYY-MM-DD` format (ISO 8601).
- `site` : name of site; general location where CAREs are set.
- `station` : name of station; a station is a location within a site; stations would typically be a few hundred meters apart and data from various stations will likely be grouped together in the analysis.
- `lat`, `lon` : location of station; if possible in decimal format, alternatively in a consistent degreesºminutes'seconds" format.
- `n_gear_in`: number of CARE put in the water at the station.
- `n_gear_ok`: number of CARE retrieved and checked to be working (light on, collector shut, etc.); this is what will be used to compute fishing effort.
- `comments`: well, comments `;-)`

Additional information such as bottom depth, weather, time of retrieval etc. can also be recorded in this table. If some information is often repeated (such as coordinates, if the stations are always the same), those can be stored in a different table and linked to this one.


A table `catches` in which information is recorded only when a fish larva is caught. Again, a minimal table would contain

```{r}
read_excel("sample_record/catches.xlsx", sheet=1, na="NA")
```

where 

- `date_out`, `site`, and `station` allow to match up with the records in `effort`.
- `family`, `genus`, `species`: taxonomic denomination of the larvae. When an element cannot be determined:
    - either confusion between several taxa is possible (several genera, or several species --usually noted spp.), in that case leave it blank
    - or the larvae can be determined as all belonging to one taxon (one genus, one species) but this taxon cannot be identified, in that case give it a generic name such as `gen1`, `sp1`, `sp_dark`, and, if possible, document it in the comments (see example above).
- `n`: number of larvae.
- `comments`: well, still comments, but on the fish now.

Additional information such as developmental stage, size, etc. can be recorded here.

Zero catches do *not* need to be recorded (and should *not* be) because they can be deduced from the matchup with `effort`. For example, one CARE was put in the water at site A on 2012-10-24 but nothing is recorded in `catches`, meaning that nothing was caught.





